### 2025-05-06 manuscript/manuscript.Rmd
Committed from clean-history branch of gag_ckmr

---
title             : "Application of Close-kin mark–recapture (CKMR) to Protogynous Fishes"
shorttitle        : "\\mbox{}"

author: 
  - name          : "Claudia Friess"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : ""
    email         : "elasmophile@gmail.com"
    role:         
  - name          : "Eric C. Anderson"
    affiliation   : "2"
    role:
  - name          : "Lisa E. Ailloud"
    affiliation   : "3"
    role:
  - name          : "David S. Portnoy"
    affiliation   : "4"
    role:
  - name          : "Susan K. Lowerre-Barbieri"
    affiliation   : "1,5"
    role:

affiliation:
  - id            : "1"
    institution   : "Florida Fish and Wildlife Research Institute, St. Petersburg, FL, USA"
  - id            : "2"
    institution   : "NOAA, National Marine Fisheries Service, Southwest Fisheries Science Center, Fisheries Ecology Division, Santa Cruz, CA, USA"
  - id            : "3"
    institution   : "NOAA, National Marine Fisheries Service, Southeast Fisheries Science Center, Charleston, SC, USA"
  - id            : "4"
    institution   : "Marine Genomics Laboratory, Texas A&M University, Corpus Christi, Corpus Christi, TX, USA"
  - id            : "5"
    institution   : "University of Florida, Fisheries and Aquatic Sciences, Gainesville, FL USA"

authornote: ""

abstract: |
  Abundance estimates for marine fish are difficult to obtain and typically rely on 
  stock assessment models that assume only females matter to productivity. Many fished 
  species, however, are sequential hermaphrodites for which fishing can drastically 
  skew sex ratios, raising concerns about sustainability, especially if managed based 
  on female-only reference points. There is interest in integrating Close-Kin Mark-Recapture 
  (CKMR) into stock assessments to reduce uncertainty in abundance estimates, but this 
  has not previously been done for sequential hermaphrodites. Here we derive the CKMR kinship probabilities 
  for protogynous hermaphrodites and include them in an integrated stock assessment framework
  for estimating abundance. Using simulation, we evaluate the robustness of this approach to 
  uncertainty in male contribution to reproductive success and genetic data availability. 
  We show that CKMR can produce unbiased abundance and sex ratio estimates but only 
  if the relative contribution of male and female age classes to reproductive success 
  are not misspecified in the estimation model and the level of mitochondrial haplotype diversity 
  in the population is well estimated. We also highlight the potential for CKMR data to estimate 
  parameters of the sex transition and male reproductive contribution functions in an 
  integrated modeling framework.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "simulation, mtDNA, reproductive resilience, integrated stock assessment, hermaphrodites, genetic tagging"

bibliography      : ["citations.bib", "r-references.bib"]
csl               : "cjfas.csl"

floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no
always_allow_html : true

classoption       : "man"
output            : papaja::apa6_pdf
header-includes   : 
  - \usepackage{color}
  - \usepackage{tipa}
  - \usepackage{wasysym}
  - \usepackage{fancyhdr}
  - \usepackage{mathrsfs}
  - \thispagestyle{empty}
  - \usepackage{hanging}
  - \captionsetup[figure]{labelfont=bf}
  - \captionsetup[table]{labelfont=bf, labelsep = period, textformat = simple}
  - \raggedbottom
  - \newcommand{\tprob}{P^{F\rightarrow M}}
  - \newcommand{\wt}[1]{\omega_#1}  % recruitment deviations
  - \newcommand{\rbar}{\bar{R}}  % mean recruitment
  - \newcommand{\fbarr}{{\bar{F}_\mathrm{rec}}}  % mean fishing mortality from rec fishery
  - \newcommand{\frecdev}[1]{fdev^\mathrm{rec}_#1}  % annual deviations from mean mortality from rec fishery
  - \newcommand{\fbarc}{{\bar{F}_\mathrm{com}}}  % mean fishing mortality from commercial fishery
  - \newcommand{\fcomdev}[1]{fdev^\mathrm{com}_#1}  % annual deviations from mean mortality from rec fishery
  - \newcommand{\Mage}[1]{\dot{M}_#1}  % age-specific instantaneous natural mortality
  - \newcommand{\vulr}[1]{\dot{v}^\mathrm{rec}_#1}  % age-specific vulnerability to recreational fishery
  - \newcommand{\vulc}[1]{\dot{v}^\mathrm{com}_#1}  % age-specific vulnerability to commercial fishery
  - \newcommand{\dFec}[1]{\dot{Fec}_#1}    % age-specific relative fecundity
  - \newcommand{\Af}[1]{\dot{Mat}_#1}    % age-specific fraction of females mature
  - \newcommand{\Pfm}[1]{\dot{P}^{\mathrm{F}\rightarrow\mathrm{M}}_#1}    % age-specific prob of transitioning from female to male
  - \newcommand{\Imp}[1]{\dot{P}^\mathrm{M}_#1}    % equilibrium proportion of males amongst age a fish
  - \newcommand{\It}[1]{\mathring{I}_{#1}}   % recruitment index in year t
  - \newcommand{\pacom}[2]{\mathring{p}^\mathrm{com}_{#1,#2}}   % relative abundance of age a females in year t
  - \newcommand{\pacr}[3]{\mathring{p}^\mathrm{ckmr}_{#1,#2,#3}}   % relative abundance of age a males in year t
  - \newcommand{\Z}[2]{Z_{#1,#2}}   % total instantaneous mortality at time t for age a fish 
  - \newcommand{\AS}[2]{S_{#1,#2}}   % total annual survival
  - \newcommand{\lxf}[1]{s_{#1}}     % survivorship
  - \newcommand{\Nf}[2]{N^\mathrm{fem}_{{#1}, {#2}}}  % number of females of age a at time t
  - \newcommand{\Nm}[2]{N^\mathrm{mal}_{{#1}, {#2}}}  % number of males of age a at time t
  - \newcommand{\K}[4]{\mathring{K}^{\mathrm{#1},{{#4}}}_{{#2},{#3}}}
  - \newcommand{\KHS}[3]{\mathring{K}^{\mathrm{HSP},{{#3}}}_{{#1},{#2}}}
  - \newcommand{\KHSs}[2]{\mathring{K}^{\mathrm{HSP},{{#2}}}_{{#1}}}
  - \newcommand{\KPOPT}[3]{\mathring{K}^{\mathrm{POP},{{#3}}}_{{#1},{#2}}}
  - \newcommand{\KHSD}[2]{\mathring{K}^{\mathrm{HSP}^{\neq _{a,b}}}_{{#1},{#2}}}   % number of half-sibling pairs (born c1 and c2) with different mtDNA 
  - \newcommand{\KHSS}[2]{\mathring{K}^{\mathrm{HSP}^{=_a}}_{{#1},{#2}}}   % number of half-sibling pairs (born c1 and c2) with same mtDNA
  - \newcommand{\KPOP}[3]{\mathring{K}^{\mathrm{POP}^{#3}}_{{#1},{#2}}}  
  - \newcommand{\KPOPF}[2]{\mathring{K}^\mathrm{POP_F}_{{#1},{#2}}}   % number of half-sibling pairs (born c1 and c2) with same mtDNA
  - \newcommand{\KPOPM}[2]{\mathring{K}^\mathrm{POP_M}_{{#1},{#2}}}   % number of half-sibling pairs (born c1 and c2) with same mtDNA 
  - \newcommand{\HSPD}{\mathrm{HSP}^{\neq}}
  - \newcommand{\HSPS}{\mathrm{HSP}^=}
  - \newcommand{\POPD}{\mathrm{POP}^{\neq}}
  - \newcommand{\POPS}{\mathrm{POP}^=}
  - \newcommand{\KNot}[2]{\mathring{K}^\mathrm{Not}_{{#1},{#2}}}     % number of pairs that are neither HSD or HSS
  - \newcommand{\KNote}[2]{\mathring{K}^{\mathrm{Not}=_a}_{{#1},{#2}}}
  - \newcommand{\KNotu}[2]{\mathring{K}^{\mathrm{Not}{\neq}_{a,b}}_{{#1},{#2}}}   
  - \newcommand{\Hmt}{H_\mathrm{mt}}  % mtDNA haplotype diversity
  - \newcommand{\Hmts}{H_\mathrm{mt}^{\mathrm{OM}}}  % simulated mtDNA haplotype diversity
  - \newcommand{\Hmte}{H_\mathrm{mt}^{\mathrm{EM}}}  % mtDNA haplotype diversity assumed in the EM
  - \newcommand{\Mat}{\mathrm{Mat}}
  - \newcommand{\Fec}{\mathrm{Fec}}
  - \DeclareMathAlphabet{\mathbbold}{U}{bbold}{m}{n}
  - \newcommand{\comment}[1]{\textcolor{red}{#1}}
  - \newcommand{\ERRO}{\mathrm{ERRO}}
  - \newcommand{\ETRO}{\mathrm{ETRO}}
  - \newcommand{\sH}{\mathscr{H}}
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
library(tidyverse)
library(stargazer)
library(kableExtra)
library(gridExtra)
library(nplyr)
library(patchwork)
library(ggh4x)
library(cowplot)
library(gridExtra)
library(grid)
library(scales)
library(here)

# caching for quick knitting.  
#    set cache = FALSE to not cache
#    set cache.extra to a new value to remove all caches and re-evaluate
knitr::opts_chunk$set(cache.extra = 4, cache = FALSE)

```

```{r figures, include = FALSE}
source(here("code", "analysis", "results_tables.R"), local = knitr::knit_global())
source(here("code", "analysis", "results_figures.R"), local = knitr::knit_global())
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# Introduction

Generating accurate estimates of marine fish population abundance is one of the most challenging 
yet essential mandates of fisheries science. To meet it, ever-more sophisticated stock assessment
models, capable of integrating a variety of data sources, are being developed [@punt2020essential].
At the core, however, these models are based on the same basic fish population dynamics 
equations developed for gonochoristic species decades ago [@beverton1993dynamics; @hilborn1992quantitative]. Similarly unchanging
has been the central overarching goal of fisheries management: optimize resource exploitation by balancing current 
extraction against future productivity, or reproductive potential. 
The key, then, is how we define, both conceptually and operationally, "reproductive potential". In terrestrial mammals, it is driven by the 
number and fecundity of the females. This paradigm has been adopted for marine fish in the form of measures
like spawning stock biomass (SSB) or total egg production, although problems with this approach are increasingly 
recognized [@sharma2019recruitment]. The reproductive resilience paradigm posits that marine fish reproductive strategies 
are unlike those seen on land and evolved to achieve lifetime reproductive success and maintain population 
stability within highly variable environments [@tringali2023reproductive; @lowerre2017reproductive]. 
These reproductive strategies include extreme fecundity, external fertilization of pelagic eggs, 
offspring survivorship affected by where and when fish spawn, and sequential hermaphroditism.
Both protandry (reproducing as males first and then as females) and protogyny (reproducing as females first and changing sex to male) 
occur in marine fishes, but protogyny is most common [@pla2022switches], especially in warmwater
systems [@lowerre2023unified].

Delayed male maturation, such as seen in protogynous species, is predicted to evolve when there
is male competition for females. Sex allocation theory predicts that adult sex ratios are driven
by processes that result in optimal mean reproductive success for the population [@charnov1982theory; @warner1988sex] and the size 
advantage model predicts that a fish will change sex when reproductive success at age is greater for
the terminal sex [@ghiselin1969evolution; @warner1975reproductive; @munoz2003new]. 
Fisheries target larger animals, truncating population age and size structure [@barnett2017old], 
which disproportionately increases the risk of overexploitation for males in protogynous species.
Yet, many protogynous species support important fisheries, 
such as the keystone predator California sheephead, *Bodianus pulcher*, [@alonzo2004status; @hamilton2015exploitation] and 
heavily-fished serranid groupers in the Gulf of Mexico (GOM) and Caribbean Sea.

Two current areas of active research in fisheries stock assessment are defining and measuring 
reproductive potential for protogynous hermaphrodites [@alonzo2008importance; @brooks2008stock;
@alonzo2005sex] and incorporating genetic data into integrated models [@punt2020essential; @punt2024including].
Of particular interest is Close-Kin Mark-Recapture (CKMR) [@nielsen2001statistical;
@julius2001allele; @bravington2016close], a method that
uses genetic data to identify kin pairs in a sample and treats those
kin pairs as "recaptures" in a mark-recapture-like framework [see @casas2023review for a 
review of CKMR and its application to fisheries].
CKMR has been used to estimate absolute abundance for a number of marine and freshwater fish 
species with a range of reproductive strategies, including the
gonochoristic southern Bluefin tuna, *Thunnus thynnus*, [@bravington2016absolute; @davies2020next], 
semelparous Chinook salmon, *Oncorhynchus tshawytscha*, [@rawding2014genetic], and elasmobranchs which have a fixed-ratio 
reproductive strategy [@hillary2018genetic; @delaval2023evaluating].
Integrating CKMR into a stock assessment requires including the CKMR pseudolikelihood
[@bravington2016close] as one component in the stock assessment model.
This CKMR pseudolikelihood relates the observed number of genetically identified closely related pairs
(e.g., parent-offspring or half-sibling-pairs) in a sample to the
number expected given the size of the population. The expected
number of kin pairs is found by deriving a set of _kinship probabilities_, each one being
the probability that a pair of samples are close kin, given the size of the population and
additional information that describes the likelihood of an individual surviving and
reproducing over time relative to other individuals with similar demographic characteristics.

To date, nearly all CKMR applications have been to gonochoristic species and have 
assumed fecundity drives reproductive success and that this success scales
equally between males and females. The life history of protogynous hermaphrodites 
complicates application of CKMR because 
sex ratios are typically skewed toward females, and, amongst half-sibling
pairs, one fish's mother could be another's father. In fact, CKMR kinship probabilities 
for protogynous hermaphrodites have not previously been derived. Additionally, it is 
unknown whether traditionally-used heuristics for the CKMR sample sizes required
to produce accurate and precise abundance estimates in gonochoristic species are 
also appropriate for protogynous hermaphrodites. Finally, even though current best
practice in stock assessments of protogynous hermaphrodites is to assume male and
female SSB contribute equally to reproductive potential [@brooks2008stock], it has not been determined 
whether this is, in fact, the case, nor do we know if and how a declining male sex ratio
affects population productivity.

In this study, we derive the kinship probabilities for protogynous hermaphrodites,
extending the work of @bravington2016close to use the frequency of kinship types 
observed in genetic data to estimate abundance and demographics in protogynous 
species. We then conduct a simulation study
to investigate the robustness of these estimators to various levels of genetic information availability
as well as uncertainty in male contribution to reproductive success and key reproductive and genetic parameters. 
The simulation consists of an individual-based operating model (OM) that generates pedigrees for a population 
of protogynous hermaphrodites experiencing increasing levels of fishing mortality 
and concurrent population decline and increased sex-ratio skew. Each simulation creates millions of
fish _in silico_ with known birth and death dates, sex, and mortality causes, and a known pedigree 
that connects them all. We then sample individuals from the OM using different data collection 
strategies—varying both the number of years and the sample size—to provide both traditional fisheries data and
CKMR kin pairs for analysis in a simple, integrated fisheries stock-assessment-style estimation 
model (EM). Finally, we quantify the bias in estimated abundance and male sex ratio 
under different OM configurations regarding male reproductive success and when key genetic and reproductive 
parameters or quantities are treated as known, estimated, or misspecified in the EM. 
Of particular interest was whether there is sufficient 
information in the CKMR data to estimate the shapes of the sex transition and male age-specific reproductive
output functions used in the CKMR likelihood.

# Methods

This section is organized as follows: 
we first describe the CKMR kinship probabilities for protogynous hermaphrodites, then
present the different components of the simulation study to test this model: the (individual-based) 
operating model (OM) to simulate 'reality', the sampling model to generate 'observed' data from 
the OM, and finally the integrated fisheries stock assessment estimation model (EM) that uses
the 'observed' data and the derived CKMR kinship probabilities to estimate a set of parameters 
that allow generating 'estimated' abundance which can then be compared to the 'true' abundance 
generated by the OM. For each simulation component, we describe the configurations or scenarios 
we used to test the performance of our CKMR estimators (Figure \@ref(fig:simoverview-plot)).

## CKMR Model

There are several close kin relationships that could be included. For simplicity, we
limit this study to half-sibling pairs (HSPs) and parent-offspring pairs (POPs) and
assume that these can be identified without error. We furthermore assume that ages of 
sampled individuals are known without error, sampling is conducted on dead
individuals, and age-specific survival is independent of sex. We assume the species is monandric (i.e., all males
originate from females) and that sex change, if it occurs, happens at the
beginning of the annual interval, followed by reproduction, then mortality/fishing/sampling. 
The full CKMR model derivation can be found in Appendix A. The probabilities of observing
HSPs (Equations A1-A7) account for the unknown age and sex of the shared parent as
well as the possibility that the first-born's mother could be the second-born's father.
Similarly, the probabilities of observing POPs (Equations A9-A13) account for the fact that
the parent, if it was male at the time of sampling, could have been female when it
gave birth to the offspring in the pair. Sex change is included via the age-specific
probability of a female changing sex to become male, $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$
(see Table \@ref(tab:table1) for this and other notation used in this paper).

The ability to identify---or at least infer---the sex of a sampled individual's parent is crucial for 
applying the model, and in our model this is accomplished by sequencing the fish's
mitochondrial DNA (mtDNA) which is inherited from the female parent. For the purposes
of the simulation analyses, we assumed that different mtDNA haplotypes were equifrequent,
so that we were able to parameterize the mtDNA haplotype frequencies simply by $H_\mathrm{mt}$---the
probability that two randomly sampled mtDNAs from the population
are different in sequence. This quantity is also commonly called the mtDNA
haplotype diversity [@avise2012molecular], and parameterizing the simulations in terms of it allowed
us to avoid simulating mtDNA sequences
in all simulated individuals and also simplified the CKMR likelihood used to analyze the simulations as described
in the Supplement. An $\Hmt$ of 1 would indicate a 100% certainty
that two HSPs with the same mtDNA sequence shared a mother, while any value <1 means there is
a chance (which equals $1 - \Hmt$) that their mtDNA sequence is the same even though they did not
share a mother. We incorporate uncertainty due to $\Hmt$ into our CKMR model
equations (Equations A8, and A14-A17) and in the simulation, we explore the effects of 
misspecifying both $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\Hmt$ in the estimation model (see below).

## Model Species

We parameterized our simulation model based on life history data for Gag, 
*Mycteroperca microlepis*, in the GOM. Gag live to be over 30 years old and 
mature around age 4. Gag exhibit clear depth preferences with life stage, sex, 
and spawning, utilizing estuarine nursery habitat as young of the year, 
moving to nearshore sites with structure as young females (~ages 1 to 4), 
and then moving offshore to spawn as they mature. Individuals undergoing sex transition occur before, 
during, and after the spawning season and in both nearshore all-female 
pre-spawning aggregations and in offshore spawning grounds [@lowerre2020testing].
Gag spawn at the shelf edge from February to mid-April, likely in pairs.
They produce pelagic eggs that settle in estuaries after 35 to 45 days [@fitzhugh2005spatial].
Males are thought to remain in deeper waters on the shelf edge year-round whereas some females return to 
nearshore structure after the spawning season. Although a male as young as four years 
old has been observed, age at 50% male is approximately 12 years old.

Gag are highly sought after in both the commercial and recreational fisheries in 
the GOM. The majority of the catch quota is allocated to the recreational 
sector, and estimates of recreational landings are considered highly uncertain [@SEDAR2022; @NOAA2023]. 
Gag have been estimated to be overfished and experiencing overfishing almost 
consistently since the late 1960’s [@SEDAR2022]. The spawning stock biomass estimated 
in the 2022 stock assessment was about 2% of unfished levels, and the sex ratio has declined from an 
estimated 32% male in unfished conditions to between 1 and 5% male in recent years 
[@lowerre2020testing; @SEDAR2022]. It is unknown to what extent sperm limitation 
is responsible for the continued decline of the gag population [@lowerre2021gag].

## Operating Model

The OM is a discrete-time, individual-based model (IBM) implemented in R based on the 'fishSim' 
package [@R-fishSim]. We modified the package functions to accommodate 
our study needs (e.g., by adding the ability to model protogynous hermaphrodite 
life histories) and rewrote the slower-running R functions in C++ using the 'Rcpp' 
package [@eddelbuettel2013seamless] to increase computational efficiency. The basic
annual sequence, modeled loosely on gag life history,
was that sex transition and then spawning occurred at the beginning
of the year, followed by mortality, CKMR sampling, and age incrementation (Figure \@ref(fig:flowchart-plot)).
For computational efficiency, the simulation starts at age 1, given that most age-0 gag
are not available to the fishery and suffer high rates of natural mortality.

Simulated reproductive value for females and males of age $a$, denoted as  $\mathrm{RV}_{a,F}$ and
$\mathrm{RV}_{a,M}$, respectively, were used to produce age-1 offspring and assign parentage.
$\mathrm{RV}_{a,F}$ represents the age-specific expected number 
of age-1 offspring produced per female. This was modeled as a power function 
(Figure \@ref(fig:RC-plot), panel A) derived 
from female gag batch fecundity-at-age data, and scaled to achieve a stable population in 
the absence of fishing. To help produce recruitment variability, $\mathrm{RV}_{a,F}$ 
was randomly scaled up or down annually by sampling a reproductive value, 
$\mathrm{RV}_{a,t,F,b}$ for each breeding group, $b$, from a $\mathrm{Uniform}(0,1)$ 
distribution and rescaling such that $\sum_a{\mathrm{RV}_{a,F}e^{\eta_t}} = \sum_a{\sum_b\mathrm{RV}_{a,t,F,b}}$.
Here, $\eta_t$ represents the annual scaling factor, ensuring that the total RV 
across all ages and breeding groups remains consistent.
The realized number of age-1 offspring in a given year $t$
for a given female of age $a$ in breeding group $b$ was a Poisson random draw from 
$\mathrm{RV}_{a,t,F,b}$. All males were considered mature, and $\mathrm{RV}_M$ 
was used as the age-specific probability of being the father
to any age-1 offspring produced by females in the male's breeding group. The true shape and 
parameter values of $\mathrm{RV}_M$ are unknown. Here, we modeled it as a power function, and in 
the base OM its exponent was set equal to that of $\mathrm{RV}_F$ (Figure \@ref(fig:RC-plot), panel A), 
which corresponds to assuming age-specific $\mathrm{RV}_M$ equals that of the females (the current
assumption of the gag stock assessment). Mating was assumed to 
be fully random within breeding groups.

Gag life history data and age schedules (Figure S1) were 
taken from the 2022 gag stock assessment [@SEDAR2022]. To simulate total
annual survival probability at age, we followed the fisheries modelling approach 
of separating survival into age-specific, instantaneous natural
and fishing mortality components, $M$ and $F$, respectively.
Fishing mortality was further separated by fishing fleet. Traditional fisheries models
use stock-recruitment relationships to represent the density-dependent effects that
result in reduced recruitment at low spawner abundance. We instead explicitly incorporated
density-dependent age-1 survival into our simulations, such that
age-1 natural mortality had density-independent as well 
as density-dependent components. The density-dependent component was assumed to 
scale linearly with juvenile (ages 1 and 2) relative densities (relative to unfished time, $t_0$) such that 
at a relative juvenile density of 1, age-1 density-dependent survival was about 
17 percent, and at a relative juvenile density of 0.1 it was about 84 percent. 
Age-1 natural mortality was furthermore subject to random variation $\delta$ to 
help produce random recruitment variability. The total probability of survival in year
$t$ for age $a$ was calculated as:
\begin{equation}
S_{a,t} =
\begin{cases}
\begin{aligned}
&\exp\biggl[-\biggl(M_a + \max(1.75 \times \frac{\sum{N_{a\in\{1,2\},t}}}{\sum{N_{a\in\{1,2\},t_0}}}  + \delta_t, 0) + \sum_{f} v_{a,f}F_{t,f}\biggr)\biggr] & \text{if } a = 1 \\
&\exp\biggl[-\biggl(M_a + \sum_{f} v_{a,f}F_{t,f}\biggr)\biggr] & \text{if } a > 1
\end{aligned}
\end{cases}
\end{equation}
where $F_{t,f}$ is the mortality due to fleet $f$ and $v_{a,f}$ is the
vulnerability of age-$a$ fish to fleet $f$. Two fleets were 
simulated: one recreational and one commercial. Vulnerability
(i.e., the combined effect of gear selectivity 
and fish availability) for the recreational fleet was modeled as a dome-shaped 
gamma function while commercial vulnerability was specified as a logistic function 
(Figure S2, panel A), meaning the oldest individuals are only selected 
for by the commercial fleet. The simulated fishing mortality time series (Figure 
S2, panel B) followed the trend estimated by the 2022 gag
stock assessment.

Probability of reproducing in a given year was described by a logistic function, 
and the sex transition function was modeled as a cumulative normal distribution function 
(Figure S1). We simulated 40 implicitly spatially-distinct breeding groups. 
Within the spawning season, mating was restricted to individuals within the same 
breeding group. All offspring were female 
at birth and entered the population at age 1 as members of their parents’ breeding 
group. Females were allowed to switch breeding groups (see Supplement for details). Males 
did not switch breeding groups but instead remained in the breeding group of their 
sex transition year until death, consistent with the view that male gag stay on the 
spawning grounds year-round while females migrate between spawning and feeding areas.

Mortality, sex transition and whether a female reproduced in a given spawning 
season were simulated by drawing random values from uniform distributions 
between zero and one. These random draws were compared to the corresponding probabilities 
(such as the probability of a female of age $a$ reproducing), with the event (e.g., reproduction) 
occurring if the random draw was less than the probability.
Mortality source was determined for each individual 
that died in a given year. The probability of having died from natural mortality, 
commercial fishing, or recreational fishing was equal to the normalized instantaneous rates
of each mortality source for a given age class and year.

The simulation was run for 119 years, starting with an unfished population and 
ending with a highly depleted population. For each OM scenario (see below), 40
independent replicates of the population were simulated, providing 40
unique whole-population pedigrees.

### OM Scenarios

We simulated four alternative versions of 'reality' in our OM to explore the
effect of uncertainty about male contribution to reproductive success: in addition to
the base case, one scenario included sperm limitation and two scenarios
simulated (age-specific) male reproductive value functions ($\mathrm{RV}_M^\mathrm{OM}$)
that differed from those of the females ($\mathrm{RV}_M^\mathrm{EM}$). In the base OM, male 
and female reproductive value were equal, corresponding 
to what is currently assumed in the gag stock assessment, but in two alternative 
scenarios, we specified $\mathrm{RV}_M^\mathrm{OM}$ functions that increased more or less steeply
from $\mathrm{RV}_F^\mathrm{OM}$ (Figure \@ref(fig:RC-plot), panel A).

The true $\mathrm{RV}_M$ in protogynous hermaphrodites, including gag,
is unknown. It is assumed that sperm limitation exists but it has not been
measured. However, it has been shown that skipped spawning is relatively common in female gag,
and this is thought to be linked to low male abundance [@lowerre2022spawning].
In the base OM configuration, declining male abundance had no effect (meaning all eggs produced
by females will get fertilized, regardless of sex ratio). Alternately, we simulated the effect of
sperm limitation as an increased
likelihood of a female not producing any offspring in a given spawning season as the sex ratio
declines. Note that a female skipping spawning was independent between years and thus there was no systematic
skipped spawning such as alternate year spawning. The relationship between the 
current proportion of males in a spawning group relative to unfished levels, $p_M$,
and the probability of females in that spawning group skipping spawning was
modeled as an exponential function of the form $1-\exp(-{\tau p_M})$ (Figure
\@ref(fig:RC-plot), panel B), where $\tau$ controls the slope of the curve. We used a sigmoid
function to map probability of spawning to age, such that skipped spawning was more likely
in younger females. Thus, skipped spawning serves as a de facto modifier of the vector 
describing the age-specific probability of a female producing offspring in a given year.

## Sampling Model
Three types of observations were generated from the OM for use in the
integrated assessment model: CKMR kinships 
<!--
(i.e., the counts $\KHSS{c_i,c_j}{g_i,s_i}$, $\KHSD{c_i,c_j}{g_i,s_i}$,
$\KPOP{c_i,c_j}{g_i,s_i}{=}$, $\KPOP{c_i,c_j}{g_i,s_i}{\neq}$, and
$\KNot{c_i,c_j}{g_i,s_i}$), 
-->
(i.e., counts of the number of observed HSPs and POPs sharing or not sharing mtDNA
as well as the number of unrelated pairs),
a recruitment (age-1) relative abundance index, and age compositions. 
These served as data inputs for the estimation model (see below). 
Note that, unlike traditional fishery models, we did
not simulate or fit to catch data; scale for absolute abundance estimates was provided solely by
the CKMR data.

Samples for CKMR were drawn
during the final years of the simulation, from all dead individuals in the population, 
sampling age 1 for juveniles and ages 4+ for adults. In reality, adult samples would be obtained
through a combination of fishery-independent and fishery-dependent sampling, as most of the
males would need to be sampled in marine protected areas, given their scarcity in open areas [@lowerre2020testing].
The kin relationship between every pair of sampled
individuals was identified from the pedigree, and identity by descent of mtDNA between relatives
was recorded from the sex of the parent (for POPs) and of the shared parent (for HSPs).  Pairs that did not share mtDNA identical-by-descent were assigned matching
mtDNA with probability $1-H_\mathrm{mt}$. These observations were tallied according
to both individuals' birth years as well as the first-born's sampling year and sex at time of sampling.
Only pairs of samples with different birth years ("cross-cohort
comparisons") were included in the CKMR likelihood to minimize the effect of
interannual and inter-individual variation in the number of offspring [@waples2022close].
The total number of CKMR samples collected across all sampling years (see below) was chosen as various multiples of the
$10\sqrt{N}$ guideline from @bravington2016close for a single sample of a
roughly equal mix of juveniles and adults. We used the average simulated abundance of
age 2+ individuals over the CKMR sampling period as $N$ in the total sample
size calculation and the average sex ratio using all ages to partition samples between males and females.
The effect of increasing or decreasing the relative proportion of males sampled is explored in the supplement.
Thus, the sample sizes varied across simulation replicates of the OM.
The total genetic samples
were spread out over the CKMR sampling period, such that if there were fewer
sampling years, more samples per year would be collected. For each of the 40
simulated pedigrees per OM configuration, dead individuals were subsampled with a different random seed ten 
times to create ten distinct CKMR data sets per pedigree, for a total of 400 CKMR data sets per
OM configuration (the subsampling effect is explored in the Supplement).

The observed recruitment index, $I_t$, was simulated as:
\begin{equation}
I_t=\frac{N_{a=1,t}}{\max_t(N_{a=1,t})}e^{\epsilon_t}
\end{equation}
where $\epsilon$ is a normally-distributed error term with lognormal bias correction:
$\epsilon_t=\mathcal{N}(0,\sigma_I)-0.5\sigma_I^2$.
$\sigma_I$ was set to 0.01, thus producing very low variability to better
assess the uncertainty due to the CKMR and hermaphroditic aspects of the model
in this study. 
The age-1 population was indexed after density-dependent and stochastic mortality but prior to density-independent
mortality to align as closely as possible with the EM.

Age compositions were recorded as the relative proportion of each age class
in any given year, where age-1 abundance, once again, reflected pre-density-independent
mortality status. Note that we did not sample the age-composition but instead assumed it
was known perfectly. As with the unrealistically low abundance index $\epsilon$, this
was done to remove uncertainty due to fisheries data and focus solely on CKMR.

### Genetic Sampling Scenarios

To evaluate the impact of sampling on the number of observed kin pairs and 
estimator performance, we specified two
different sampling periods and three different sample sizes. In the base case, 
CKMR samples were collected over three years. Ten years
of sampling was explored as an alternative to see whether the time period over which sampling
is spread out affects the number of related pairs found and EM performance.
The base CKMR sample size, $n_\mathrm{base}$ was equal to the $10\sqrt{N}$
heuristic. The EM was additionally run with 50% and 150% of that base sample size.

## Estimation Model

We employed a standard age-structured deterministic fishery population dynamics
model and integrated the CKMR estimators derived in the Appendix to see how well the 
simulated sex-specific abundances could be estimated, given the data generated from the OM. 
Assumed known were all life history and fishery
vulnerability age schedules except sex change probability, $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$,
and age-specific male reproductive value, $\mathrm{RV}_{a,M}$. The inflection point of the two-parameter 
cumulative normal distribution $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$
function was always assumed known but the standard
deviation, $\mathrm{STF}_{SD}$, which controls the function's shape was estimated in some EM runs. 
The exponent of the $\mathrm{RV}$
power function was assumed known in the EM only for females 
($\mathrm{RV}_F^\mathrm{OM} = \mathrm{RV}_F^\mathrm{EM}$, more on this below). Additional estimated parameters were mean
recruitment $\rbar$ and fleet-specific fishing mortality rates $F$ as well as annual
deviations for each. No stock-recruitment relationship was simulated or attempted
to be estimated, and unfished abundance was also not estimated since the focus of
the study was to estimate scale (i.e., absolute abundance) and sex ratio in recent years. Full equations
for the total objective function and its components are presented in Appendix B. 
The estimation model was implemented in AD Model Builder [@fournier2012ad].

### EM Scenarios

The EM for the base case included the full CKMR pseudolikelihood (for both HSPs and POPs)
and assumed that the degree of mitochondrial haplotype diversity in the population ($\Hmt$),
$\tprob_{a}$, and $\mathrm{RV}_M$ were known. 
Alternative scenarios tested models that included HSPs only in the likelihood or
contained misspecification (Figure \@ref(fig:simoverview-plot)). 
One misspecification was an incorrect $\Hmt$, which affects our ability
to correctly infer the sex of an unobserved parent or of an observed parent at a time
prior to sampling, potentially biasing CKMR kinship probabilities.
Other misspecifications were that skipped spawning was simulated 
in the OM but not accounted for in the EM and that 
$\mathrm{RV}_M^\mathrm{EM} = \mathrm{RV}_F^\mathrm{EM}$ when 
$\mathrm{RV}_M^\mathrm{OM} \neq \mathrm{RV}_F^\mathrm{OM}$. 
Finally, we allowed the EM to estimate $\mathrm{RV}_M$ to 
assess whether the model could infer this unknown quantity from the CKMR data, 
and to estimate $\tprob_{a}$ to determine if the CKMR data could allow estimating the sex ratio.

In evaluating results when HSPs only are included in the likelihood, we are
asking whether our ability to estimate abundance deteriorates when
POPs are excluded from the CKMR pseudolikelihood. In reality, POPs 
in open marine fish populations, even severely depleted ones, are going to be very rare 
and thus may not contribute much information to the model, 
so we tested the value of including POPs.

Mitochondrial haplotype diversity needs to be estimated 
outside the integrated assessment model, and we tested the effect of 
getting this wrong. While $\Hmt$ affects both the POP and HSP calculations, we focus
here on HSPs for illustrating its importance. Given that HSPs sharing a
mother are going to be much more rare in protogynous hermaphrodites than HSPs sharing a father,
having the wrong information about $\Hmt$ could incorrectly inflate or deflate
the true number of HSPs sharing a mother in the sample, thus producing biased abundance estimates.
In the base configuration, $\Hmt$ in both EM and OM ($\Hmte$ and $\Hmts$, respectively)
was set to 1. In two alternative configurations we first set $\Hmts$ higher than $\Hmte$
(1 vs 0.9) and then *vice versa*. In the case of $\Hmts$ < $\Hmte$, the observed number of HSPs with the same mtDNA ($\HSPS$) 
will be larger than the true number of HSPs sharing a mother but the EM won't know this, 
meaning it will see more HSPs sharing a mother than there truly are. This should lead to underestimating
the number of females and overestimating the number of males.\footnote{Note that $\HSPS$ is only 
synonymous with maternal HSPs when $\Hmt$ is 1, and $\HSPD$ could be either paternal HSPs or HSPs where
the shared parent changed sex.} Conversely, in the case of $\Hmte$ < $\Hmts$, the observed number 
of $\HSPS$ will be closer to the true number of HSPs sharing
a mother, but the EM will assume some of these should be HSPs with different mitochondrial DNA ($\HSPD$), 
meaning it will discount the true number of HSPs sharing a mother and overestimate female
abundance.

For the OM scenarios that simulated $\mathrm{RV}_M^\mathrm{OM} \neq \mathrm{RV}_F^\mathrm{OM}$,
if the EM assumes that $\mathrm{RV}_M^\mathrm{EM} = \mathrm{RV}_F^\mathrm{EM}$
but $\mathrm{RV}_M^\mathrm{OM}$ is steeper than $\mathrm{RV}_F^\mathrm{OM}$, 
this would mean that in the simulated population, younger males 
contribute fewer offspring and more HSPs sharing a father should be found than what the EM expects. 
This would signal to the EM that there are
fewer males in the population and male abundance should be underestimated. When $\mathrm{RV}_M^\mathrm{OM}$
is flatter (i.e., more age-invariant) than $\mathrm{RV}_M^\mathrm{EM}$, the opposite would be the case: younger males would
contribute almost as many offspring as older males, and so the probability of finding HSPs
sharing a father decreases, causing an overestimation of male abundance.

### Model Evaluation
For every assessment year $t$ and simulated observed data set $k$, we calculated percent relative error ($\mathrm{PRE}$)
between simulated and estimated sex-specific 
adult abundance and sex ratio as:
\begin{equation}
\mathrm{PRE}_{t,k} = \frac{(\mathrm{EV}_{t,k}-\mathrm{SV}_{t,k})}{\mathrm{SV}_{t,k}}\times 100,
\end{equation}
where $\mathrm{EV}$ = estimated value and $\mathrm{SV}$ = simulated value. For every $k$, 
the median PRE was calculated across years, and these medians were used to calculate median and interquartile
range ($\mathrm{IQR}$) of PRE to compare accuracy and precision of estimates, respectively, by tested scenario. 
We also calculated $\mathrm{PRE}$ for the main model parameters for every $k$. Sex ratio was
defined as the number of males divided by the total number of 3+ year-olds. We include calculation 
and discussion of an additional metric, the normalized root mean squared error, in the Supplement.

# Results

## Simulated Population
At the beginning of the simulation, the median number of adults across all OM
pedigrees was about 690,000 adult individuals (ages 3+) and about 27% of them were male.
In the final year, the median abundance was ~230,000 individuals without 
the productivity-loss feedback loop from sperm limitation
and ~100,000 with sperm limitation, and about 3% of them were male. When sperm limitation
was present the population trajectory started diverging from the scenarios without
sperm limitation in the final 35-40 years of the simulation (Figure \@ref(fig:simpop-plot)).

## Simulated CKMR Samples
The number of females sampled was about two magnitudes higher than the number of
males sampled (Figure S3). Additionally, slightly fewer overall 
female CKMR samples were collected when CKMR sampling was spread out over the longer period.
This is because 10 years of sampling included years of lower abundance that lowered
the average abundance on which CKMR sample size was based. Conversely, the number
of males sampled tended to be slightly higher for the 10-year sampling time period, because
the sex ratio used to partition sampling by sex was higher for the longer time period.
The median number of males sampled for the base OM scenario ranged from 15
individuals ($\mathrm{IQR} = 3$) when CKMR sample size was 50% of $n_\mathrm{base}$ and sampling was
conducted over three years to 50 individuals ($\mathrm{IQR} = 10$) when CKMR sample size was
150% of $n_\mathrm{base}$ and samples were collected over ten years.
For adult females, the number sampled ranged from a median of 1480 ($\mathrm{IQR} = 100$) for
50% of $n_\mathrm{base}$ and ten sampling years to a median of 4664 ($\mathrm{IQR} = 497$) for
150% of $n_\mathrm{base}$ and three sampling years (Figure S3).

The number of CKMR pairs found was higher when CKMR sampling was conducted
over a shorter time period (Figure S4). Additionally, the number
of HSPs with different mtDNA ($\HSPD$) was about an order of magnitude higher than
the number of HSPs sharing mtDNA ($\HSPS$). The median number of pairs found for $\HSPD$ ranged
from 166 ($\mathrm{IQR} = 52$) for low CKMR sample size to to 1,481 ($\mathrm{IQR} = 440$) for
high CKMR sample size (when averaged across number of sampling years),
compared to a median of 10 ($\mathrm{IQR} = 6$) to 83 ($\mathrm{IQR} = 34$) pairs for
$\HSPS$ (Figure S4). The HSP pattern of observed frequencies by
mtDNA sharing was reversed for POPs. The number of POPs sharing mtDNA ($\POPS$)
was only slightly lower than the number of $\HSPS$ pairs (a median of 7 for low CKMR sampling
to 59 for high CKMR sampling) while the number of POPs with different mtDNA ($\POPD$)
was more than two orders of magnitude lower than $\HSPD$ pairs and about six times lower
than the number of $\POPS$ pairs found (a median of 1 for low CKMR sampling
to 10 for high CKMR sampling) (Figure S4).

Some differences in the number of pairs found were apparent across OM scenarios
(Figure \@ref(fig:mainckmrpairs-plot)). First, the number of unrelated pairs was much
smaller for skipped spawning than the other scenarios while
the number of related pairs was not. This is because abundance was lower for this
scenario and so fewer individuals needed to be sampled to obtain similar numbers of
related pairs. Additionally, the number of $\HSPS$
was higher for the skipped spawning scenario because younger females
were more likely to skip spawning than older ones, meaning comparatively fewer
females contributed to the offspring produced and the likelihood of finding
HSPs sharing a mother increased.
The next major effect was that of $\mathrm{RV}_M^\mathrm{OM}$ on $\HSPD$.
When $\mathrm{RV}_M^\mathrm{OM}$ increased more steeply with age, as would be
expected based on the size advantage model, the number of $\HSPD$ more than doubled
compared to the base case.
The explanation is analogous to finding more $\HSPS$ when there is skipped spawning. 
If $\mathrm{RV}_M^\mathrm{OM}$ is steeper,
older males (of which there are fewer than younger males) are much more successful at fathering offspring and
hence the chance of finding $\HSPD$ increases.
The opposite is seen when $\mathrm{RV}_M^\mathrm{OM}$ varies little with age. Here,
we only found about two-thirds the number of $\HSPD$ compared to the base case
(Figure \@ref(fig:mainckmrpairs-plot))
because now younger males are also very successful and the offspring are more
evenly spread out among fathers, reducing the likelihood of encountering HSPs sharing
a father.

## Estimation Model Convergence

The estimation model was fitted to 67,200 different data sets: 40 pedigrees subsampled ten times each
for seven OM+EM scenarios, four parameter estimation combinations, three CKMR sample sizes, 
and two CKMR sampling intervals \footnote{plus another 16,000 data sets for additional runs we
conducted and discuss in the Supplement}. All in all, ADMB
failed to converge, after fitting the model up to three times with different data
weights, for  2,792  runs (3.4%). Forty percent of the failed runs were for the smallest CKMR sample size
runs while only 25% were for the largest sample size. 
An additional 4,722 EM runs (5.7%) were excluded because
parameters were estimated at bounds. Most of these were due
to the $\mathrm{RV}_M^\mathrm{EM}$ function exponent parameter. Finally, 311 additional runs (0.4%) were excluded
because the maximum parameter gradient exceeded 0.1.
Overall, 90.6% of the simulated data sets produced model fits
that were used in subsequent analyses. Sixteen percent of runs converged and passed
the data weight checks after the first time fitting the model and an additional
30% after the second time. If the model ran 3 times, we used the
run with the lowest negative log likelihood amongst those 3 runs.

## Effect of CKMR Information Availability on Estimation Performance

For this section, we discuss only the base OM scenario, where there was no skipped spawning
and $\mathrm{RV}_M^\mathrm{OM} = \mathrm{RV}_F^\mathrm{OM}$.

### CKMR sample size and number of sampling years

When $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\mathrm{RV}_M$ were assumed known, median abundance estimates
had no (median $\mathrm{PRE} < \pm 1\%$) or low (median $\mathrm{PRE} < \pm 2\%$) bias, with no difference between
male and female estimates. Precision was relatively high, with $\mathrm{IQR_{PRE}}$
ranging from 5% to 13% across all tested CKMR sample sizes and numbers of CKMR sampling 
years (Figure \@ref(fig:ckmrsampling-plot)). There was no difference
in median bias between three and ten years of CKMR sampling and precision
was only slightly lower for ten years of sampling. Increasing the CKMR sample size increased
precision in $\mathrm{PRE}$ by a relatively small amount (28%) when sample
size was increased from 100% to 150% of $n_\mathrm{base}$ and more significantly
(${\approx}67\%$) when the sample size increased from 50% to 100% of $n_\mathrm{base}$.

When $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and/or $\mathrm{RV}_M$ were free to be estimated, median bias remained low
(median $\mathrm{PRE_{sc}}$ ranging between -3% to 7%, mean = 1% across CKMR sampling scenarios)
but precision decreased notably ($\mathrm{IQR_{PRE_{sc}}}$ ranging from 5% to 51%, mean = 24%).
$\mathrm{IQR_{PRE_{sc}}}$ was similar between males and females when $\mathrm{RV}_M$ was estimated
(Figure \@ref(fig:ckmrsampling-plot), panels C and D) but smaller for males than females
when only $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ was estimated (Figure \@ref(fig:ckmrsampling-plot), panel B). The number
of sampling years continued having only a minor impact on estimation. For all but males
when both $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\mathrm{RV}_M$ were estimated, $\mathrm{IQR_{PRE_{sc}}}$ increased when
sampling was spread out over ten years. The largest increase (by ${\approx}11\%$) was observed for
females at 50% of $n_\mathrm{base}$ when $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ was estimated
(Figure \@ref(fig:ckmrsampling-plot), panels B and D). CKMR sample size continued to
impact precision more than the median of the abundance estimates. $\mathrm{IQR_{PRE}}$ increased from 29% to 41% when sample
size decreased from 150 to 100% $n_{base}$ and from 69% to 90% when sample size dropped from 100% to 50% of $n_\mathrm{base}$.

For the remainder of the results, we discuss only the base CKMR sampling case,
meaning $10\sqrt{N}$ samples collected over three years.

### Types of CKMR pairs used in the likelihood

Male and female abundance were estimated very well in the base OM and EM configuration.
(Figure \@ref(fig:ckmrnll-plot), panel A). Estimates for both sexes had negligible bias
(median $\mathrm{PRE} < \pm 2\%$) regardless of whether $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\mathrm{RV}_M$ were estimated or
fixed. Precision, however, was notably higher when both these functions were assumed
known ($\mathrm{IQR_{PRE}} {\approx} 7\%$). Precision was lowest for males when both functions were estimated
($\mathrm{IQR_{PRE}} = 33\%$)(Figure \@ref(fig:ckmrnll-plot), panel A). There was no difference in
abundance estimates for either sex when POPs were omitted
from the CKMR likelihood, regardless of whether $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\mathrm{RV}_M$ were fixed
or estimated in the model (Figure \@ref(fig:ckmrnll-plot), panels A and B).
To see whether this finding would be different when there is
model misspecifiation, we compared EM runs with POPs and HSPs to runs with HSPs only for the skipped spawning
scenario and found no difference in this case either (Figure S5).

### Uncertainty in mitochondrial DNA haplotype diversity

Misspecifying $\Hmt$ in the EM had major effects on abundance estimation
performance, but only when $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\mathrm{RV}_M$ were not both fixed at the true value.
When one or both of these parameters were allowed to be estimated, the
resulting biases were in the direction we predicted for the females (overestimation when $\Hmte$ < $\Hmts$
and underestimation when $\Hmts$ < $\Hmte$) but, for the males, results
were more nuanced. When only $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ remained fixed at the true value, male
bias mirrored that of the females (Figure \@ref(fig:ckmrnll-plot), panels C & D)
resulting in unbiased sex ratio estimates (Figure \@ref(fig:sexratio-plot), panels C & D)
despite high bias in absolute abundance estimates, highlighting the near-total
control the sex transition function in the EM has over estimated
sex ratios and the importance of $\HSPS$ for protogynous hermaphrodites. Allowing
$P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ to be estimated when $\Hmt$ was misspecified in the EM resulted in highly
biased sex ratio estimates that became slightly less biased when $\mathrm{RV}_M$ was also
allowed to be estimated (Figure \@ref(fig:sexratio-plot), panels C & D).

When $\Hmts < 1$, the observed number of $\HSPS$ was notably higher than when $\Hmts$ = 1
(Figure S6). Thus when the EM sees more $\HSPS$ than it expects
in the case of $\Hmts$ < $\Hmte$,
it will attempt to reduce the number of expected females to match what it sees in the genetic
data. It does this by underestimating mean recruitment (Figure S7)
along with changing the shape of $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\mathrm{RV}_M^\mathrm{EM}$,
if allowed (Figure \@ref(fig:estfuns-plot)).
When $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ is fixed, thereby locking in the sex ratio, the EM can only reduce the expected
number of females by also lowering the expected number of males, and it does this by
flattening the estimated $\mathrm{RV}_M$ function (Figure \@ref(fig:estfuns-plot)). This also reduces
the expected number of $\HSPD$, in contradiction to what the genetic data for the
males are indicating. When allowed, the EM will flatten the estimated $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ which
results in females turning into males at a younger age, reducing the bias in $\mathrm{RV}_M^\mathrm{EM}$
required to reconcile observed and expected numbers
(Figure \@ref(fig:estfuns-plot)). The opposite is true when the EM sees fewer
$\HSPS$ than it expects in the data for the $\Hmte$ < $\Hmts$ scenario: recruitment is now
overestimated (Figure S7) and the estimated
$P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\mathrm{RV}_M$ functions are steeper to create
more expected females (Figure \@ref(fig:estfuns-plot)).

## Effect of Uncertainty about Male Contribution to Reproductive Success

Generally for all OM and EM configurations,
abundance estimate errors were similar between the sexes when $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ was assumed known,
resulting in fairly unbiased sex ratio estimates (Figure \@ref(fig:sexratio-plot)). We, therefore, focus
between-sex comparisons in this section on the case when $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ was allowed to be estimated.

### Skipped spawning
When skipped spawning was simulated but not accounted for in the OM, where younger females were more likely to skip spawning
than older ones, the effect was similar in direction to setting $\Hmts < \Hmte$. Fewer females are
contributing offspring, effectively shifting $\mathrm{RV}_F^\mathrm{OM}$ so the chance of encountering $\HSPS$
increases against the EM's expectation and female abundance is underestimated (Figure \@ref(fig:malerole-plot), panel B).
Under a fixed $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$, the negative bias for both sexes was
low (median $\mathrm{PRE} < 5\%$) when $\mathrm{RV}_M$ was fixed at the true value but increased
when $\mathrm{RV}_M$ was estimated (median $\mathrm{PRE} <20\%$).
When $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ was estimated, it was slightly too flat
(Figure \@ref(fig:estfuns-plot)), the female negative bias increased to a median value of
${\approx} 28\% \mathrm{PRE}$) but male bias decreased (median $\mathrm{PRE} <5\%$).
This is because a flatter $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ would turn more young females into males, thereby increasing
the number of males estimated. Consequently, male sex ratios when skipped spawning was simulated
were positively biased when $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ was allowed to be estimated
(fixed $\mathrm{RV}_M^\mathrm{EM}$: $\mathrm{PRE} = 38\%$, $\mathrm{IQR_{PRE}} = 30\%$;
estimated $\mathrm{RV}_M^\mathrm{EM}$: $\mathrm{PRE} = 34\%$, $\mathrm{IQR_{PRE}} = 49\%$;
Figure \@ref(fig:sexratio-plot), panel B).

### Differing simulated reproductive value between the sexes
When $\mathrm{RV}_M^\mathrm{EM}$ was incorrectly assumed to equal $\mathrm{RV}_F^\mathrm{EM}$, bias in male abundance
estimates was as predicted: positive when simulated $\mathrm{RV}_M$ was flatter and negative when
it was steeper than $\mathrm{RV}_F^\mathrm{OM}$ (Figure \@ref(fig:malerole-plot)C,D). Female bias
matched that of the males when $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ was fixed, but when $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ was allowed to be
estimated, median female abundance estimates were unbiased, driven by compensation
for any erroneous $\mathrm{RV}_M^\mathrm{EM}$ in the estimated shape of $P_a^{\mathrm{F}\rightarrow\mathrm{M}}$
(Figures \@ref(fig:estfuns-plot) and S8). When $\mathrm{RV}_M^\mathrm{EM}$ was
allowed to be estimated, it was estimated fairly well, especially for the steeper
$\mathrm{RV}_M^\mathrm{OM}$ scenario, and male abundance estimates became much less biased
($\mathrm{PRE}$ reduced to ${\approx}{5}\%$ from ${\approx}{55}\%$ for flatter $\mathrm{RV}_M^\mathrm{OM}$ and reduced to
${\approx}{-6}\%$ from ${\approx}{-64}\%$ for steeper $\mathrm{RV}_M^\mathrm{OM}$; Figure \@ref(fig:malerole-plot)C,D).
Hence, sex ratios were most biased when
the transition function was allowed to be estimated but $\mathrm{RV}_M$ continued to be
misspecified ($\mathrm{PRE} = 51\%$ and $\mathrm{IQR_{PRE}} = 37\%$ for flatter $\mathrm{RV}_M^\mathrm{OM}$;
$\mathrm{PRE} = -65\%$ and $\mathrm{IQR_{PRE}} = 11\%$ for steeper $\mathrm{RV}_M^\mathrm{OM}$; Figure \@ref(fig:sexratio-plot)E,F).

## Ability to estimate sex transition and male reproductive value
The sex transition function shape was overall well estimated with little variability in
estimated shapes. The $\mathrm{RV}_M$ function was much more variable (Figures \@ref(fig:estfuns-plot) and S8).
The median was estimated fairly well except when $\Hmt$ was misspecified in the EM
but $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ remained fixed and when $\mathrm{RV}_M^\mathrm{OM}$ was close to age-invariant.
Generally, the median estimated $\mathrm{RV}_M$ shapes matched the simulated curvatures,
with the notable exception of when $\Hmts < \Hmte$.
Even so, in almost all scenarios, some runs estimated incorrect curvatures (i.e., estimated
$\mathrm{RV}_M$ increased at an increasing rate with age when it was simulated
to increase at a decreasing rate with age, and *vice versa*).

The standard deviation of $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$
($\mathrm{STF}_{SD}$) and $\mathrm{RV}_M^\mathrm{EM}$ were
strongly positively correlated, meaning as the estimated $P_a^{\mathrm{F}\rightarrow\mathrm{M}}$ function
became flatter, $\mathrm{RV}_M^\mathrm{EM}$ became steeper. This correlation increased
with increasing CKMR sampling in all scenarios except when $\mathrm{RV}_M^\mathrm{OM}$
was close to age-invariant (Figure S9).
$\mathrm{STF}_{SD}$ was moderately negatively correlated with mean recruitment
but this correlation decreased with increasing CKMR sampling (Figure \@ref(fig:estfuns-plot)).

# Discussion

This study is the first, to our knowledge, to investigate the application of CKMR to a protogynous hermaphrodite 
in an integrated assessment framework. By using a data-generating model that differed
in structure from the estimation model, we were able to confirm that the CKMR estimators
we developed were correct when the base assumptions of the data-generating model were met.
From that point, one can systematically start violating these assumptions to explore
which of them, in isolation or combined, have the largest impact on estimation performance.
This exercise underscores the importance of using simulation to evaluate the effectiveness 
of a CKMR study prior to implementation, and identify research priorities to resolve 
uncertainties most likely to obscure or invalidate results.

We showed that sex-specific abundance for protogynous hermaphrodites can be estimated in
an integrated stock assessment model and that even key parameters used to calculate ERRO, 
such as the shape of the age-specific sex transition function and individual male reproductive 
value-at-age, are estimable using genetic data, but only as long as there are no 
misspecifications in the
estimation model that bias the expected probabilities of observing CKMR kin pairs. One such
misspecification we explored is the level of mitochondrial haplotype diversity in the population
which describes the probability that two randomly sampled mtDNA sequences are different.
Even though this quantity can be easily and accurately estimated by typing the mtDNA of
a large sample of individuals,
our results underscore the importance of estimating mtDNA haplotype diversity well in order to be able to
estimate $P_a^{\mathrm{F}\rightarrow\mathrm{M}}$ and $\mathrm{RV}_M$. Previous studies
have estimated $\Hmt$ for gag on the West Florida Shelf to be ~0.84-0.98 [@jue2015shelf]. In this study, 
sample size was quite low ($N=83$) and the sequencing targeted between 562 and 781 bp of the D-loop
portion of the mtDNA molecule, with the variation in sequence length and in $\Hmt$ reflecting whether
or not indels were included in the diversity measure. If gag mtDNA haplotypes are not in fact, as we assumed,
equifrequent, haplotype frequencies will need to be estimated, incorporated into the EM as data, and
accounted for in the likelihood. The implications for CKMR estimators of rare and common haplotypes 
in the population could be explored in future simulation studies.

Age-specific reproductive contribution is a key quantity determining the outcome of stock assessment
and CKMR models, and our simulations highlight this. For gonochoristic species, the only 
reproduction-related age-schedule deemed to be of importance is female fecundity (often calculated as 
maturity times weight at age and otherwise based on batch fecundity estimates). For sequential
hermaphrodites, it is common practice to assume that reproductive value scales identically
between males and females, although there is no evidence to support this assumption. In fact,
the size advantage model of sex allocation suggests the opposite [@warner1988sex]. In our simulations we
included two scenarios with shapes for male reproductive value at age that differed from that of the
females.  When the EM falsely assumed $\mathrm{RV}_M = \mathrm{RV}_F$, this resulted in significant estimation bias, but
we were encouraged to see that there was enough information in the data to allow $\mathrm{RV}_M$ to be estimated with no or little bias
when the sex transition
function was known. This is good news because data exist to estimate $P_a^{\mathrm{F}\rightarrow\mathrm{M}}$
outside the stock assessment model. This is, in fact, what is usually done [@SEDAR2022], even though 
those estimates are also uncertain [@lowerre2021gaggrouper]. By contrast, it is unclear how one would 
go about attempting to measure the age specific elements $\mathrm{RV}_{a,M}$ of the $\mathrm{RV}_{M}$
vector in the gag population. Theoretically, one could obtain 
$\mathrm{RV}_{a,M}$ by dividing the sample age composition of adult males by the sample age composition of 
male parents which allows separating $\mathrm{RV}_{a}$ from selectivity-at-age [@davies2017advice], but this 
would be extremely difficult, in practice, for gag due to the expectation of very low numbers 
of POPs sharing a father in real-world sampling.

In our simulations, excluding POPs had no effect on estimation performance and we think this is
mainly because there were so few of them. 
As long as there are not sections of the adult population systematically and unknowingly not 
contributing offspring, census population size can be estimated using either POPs or HSPs, 
but, as has been pointed out elsewhere [@waples2022close], bias can arise for either kinship type when
systematic unknown reproductive patterns and systematic sampling bias interact. There is potential
to estimate effective breeding size rather than census size from POPs if non-breeders are not 
sampled, and from HSPs if same-cohort comparisons are included [@waples2022close]. Because HSPs were the
most important data component, increasing the proportion of males in the adult CKMR sample had no effect
on our estimator. Estimation performance was even unaffected when sampling no males at all 
(Figures S15 and S16).

The mechanism by which extremely low sex ratios in protogynous hermaphrodites might compromise
reproductive resilience is unknown. In our simulation model, we assumed it would result in some females 
(predominantly the younger ones) having no reproductive success (akin to right-shifting age at maturity), 
while that of most larger, older fish is unchanged. This is analogous to time-varying $\mathrm{RV}_{a,F}$, 
shifting toward a steeper function of age as the sex ratio
declines. Time-varying life history parameters can be included in stock assessment models,
but often the data to allow their estimation are not available.

As with all simulation studies, there are known and unknown unknowns and when combined, these
could result in unexpected outcomes, conceivably even reversing the direction of observed effects. 
Our simulation model, although fairly complex, did not account for all factors believed 
to be important to gag population dynamics. The most important of these are probably spatial effects. 
@conn2020robustness showed that CKMR estimators can be biased when there is limited dispersal and 
variation in spatial sampling probabilities. Assessing biases in spatial sampling in marine fish is
challenging given that drivers of spawning site selection, fidelity, and dispersal are rarely known. 
Gag have discrete offshore spawning sites, with assumed 
male site fidelity, pelagic early life dispersal, and estuarine nursery grounds. Early life dispersal pathways 
vary temporally, being driven by loop-current-associated remote forcing, and this affects 
recruitment success [@weisberg2014gag] and likely leads to spatial clustering of related individuals in
estuarine nursery sites. Excluding same-cohort comparisons could help mitigate this limited mixing of juveniles in nursery areas,
but the extent of this effect depends on the degree of year-to-year variability in dispersal pathways, which
is currently unknown. Gag are fished in all areas they occur, except for protected spawning grounds, but the
various fishing fleets would need to be sampled appropriately in order to counter the spatial biases
in fishing and not violate the CKMR assumption of equal sampling probability for all animals.
Future simulations could address these issues. 
An additional factor our simulation model did not consider but could be included in future studies 
is variability of size-at-age and consequent size-based effects on individual reproductive output for 
individuals of the same age. Furthermore, we assumed perfect
fisheries data, with a highly accurate and precise index of juvenile abundance and perfect, though
combined-sex, composition data. In reality, the age-0 gag index is composed of multiple data streams 
whose spatio-temporal footprints have shifted through time, making the index prone to bias and uncertainty
[@ingram2021juvenile]. And composition data are highly uncertain and particularly sparse for the private
recreational fleet [@SEDAR2022], which comprises the majority of the catch.
The effect of uncertainty in fisheries data can also be examined in future simulation studies.

We made simplifying assumptions for our CKMR estimators that will need to be addressed in any
practical applications, specifically the ability to age sampled fish without error and to perfectly 
distinguish HSPs from other kinship types. Aging imprecision is expected to increase uncertainty
within assessments, and known age-error matrices can be incorporated and integrated over in 
the CKMR likelihood.
Using the marker sets of 1000 to 3000 single nucleotide polymorphisms 
commonly applied in CKMR today, it is impossible to genetically distinguish between HSPs and
either grandparent-grandchild or full avuncular (uncle/aunt-neice/nephew) pairs, because pairs with these
relationships all are expected to share the same fraction of genome with 0, 1, or 2 gene copies
identical by descent [@thompson1975estimation]. Furthermore it can be difficult
to distinguish HSPs from less-related kin types such as half-avuncular. Though it is relatively 
straightforward to account for the presence of such non-target kin pairs in practice, e.g., by
calculating and including kinship probabilities for grandparents and avuncular pairs in the CKMR likelihood 
and/or setting one's false negative rate to ensure distinction between HSPs and half-avunculars,
we did not simulate such a process.

Our estimation model was, by necessity, much simpler than the assessment model used for gag in reality. 
Any real-world model would likely require
hand-tuning that could not be implemented in a simulation framework attempting to fit thousands of
models. It remains to be seen whether $P_a^{\mathrm{F}\rightarrow\mathrm{M}}$ and $\mathrm{RV}_M$ could actually
be estimated in the real world assessment model that includes many more parameters and data sources
than our simplistic model, but perhaps with the right data, this is an achievable future goal. Unlike
most real world fisheries stock assessments, we did not fit our assessment model to catch data. This
was done because we were interested to see how well the CKMR data alone would do in estimating scale
and because catch data for gag are highly uncertain, mainly due to uncertainty in estimating recreational
catches which make up the majority of gag removals in the GOM. It is conceivable that the CKMR data
and catch data provide conflicting information about scale and trends, similar to what is often seen
for composition and index data, that would complicate model fitting; this scenario could be explored
in future simulations. We note that many species for which management advice is needed in the
Southestern U.S. and Caribbean are data-poor and for some [e.g., Goliath grouper, *Epinephelus itajara*, 
@porch2006catch], catch data are missing entirely. For these, the prospect of incorporating CKMR into catch-free assessment models is highly promising.

\newpage

## Acknowledgements

We sincerely thank Paul Conn for his thoughtful and thorough review, 
which greatly improved this manuscript. We would like to thank Rob Ahrens whose
ADMB model code provided as part of his excellent fishery stock assessment class
served as the basis for the assessment model developed in this paper.
A portion of the writing and code development for this paper occurred while
CF and ECA were scientists-in-residence with Matthew G.~DeSaix at the
mobile High Altitude Venue for Ecological Analysis, Genetics and
Statistics on location in Moab, Utah
during April 2023.  This is contribution mHAVEAGAS-004.

## Competing interests

The authors declare there are no competing interests.

## Author contributions

C.F.: Conceptualization, methodology, software, formal analysis, validation, visualization, writing - original draft; 
E.C.A.: Conceptualization, methodology, software, formal analysis, validation, visualization, writing - original draft;
L.E.A.: Resources, methodology, writing - review & editing;
D.S.P.: Conceptualization, writing - review & editing;
S.K.L.B.: Conceptualization, methodology, supervision, funding acquisition, project administration, writing - original draft

## Funding

This research was supported by the National Oceanic and Atmospheric Administration 
(NOAA; MARFIN grant numbers NA21NMF4330504 and NA19NMF4540066)
and the U.S. Fish and Wildlife Service Sport Fish Restoration Program 
(Grant numbers F-22AF00983_00, F23AF00721-00, and F24AF00716-00).

## Data availability

All code associated with analysis is available on https://github.com/cfriess/gag_ckmr.

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::

\newpage

## Tables

```{r table1, echo=FALSE}
kableExtra::kable(table1, 
                  escape = FALSE, 
                  booktabs = TRUE, 
                  caption = "Notation and definitions used in this study", 
                  col.names = NULL,
                  longtable = TRUE) %>%
  column_spec(1, width = "6em") %>%
  column_spec(2, width = "33em") %>%
  #row_spec(0, bold=TRUE) %>%
  kable_styling(latex_options = c("hold_position", "repeat_header"), 
                position = "left", repeat_header_method = "replace")

```

(ref:simoverview-plot-fig-caption) Schematic overview of the simulation routine. The
operating model (OM) concerns the simulation of individuals in the population, with four
different scenarios listed. Observed kin pair data are obtained by sampling (sampling model) from the simulated population with different scenarios varying by the number of CKMR sampling years and 
number of samples collected. Finally, several estimation model (EM) configurations were specified. Estimated abundance and sex ratio 
were compared to simulated values by calculating percent relative error. The scenarios in each submodel highlighted by bold font are the base case.
RV = reproductive value, P(\male $\rightarrow$ \female) = sex transition probability. $\Hmts$ = simulated mitochondrial 
haplotype diversity, $\Hmte$ = mitochondrial haplotype diversity assumed in the estimation model.

(ref:flowchart-plot-fig-caption) Overview of annual steps simulated in the individual-based operating model.

(ref:RC-plot-fig-caption) Simulated A) reproductive value (RC) at age and B) probability of skipping spawning as a function of sex ratio. In A), the dashed line corresponds to relative female fecundity at age. The solid lines are male RC, where red is the OM scenario where $\mathrm{RV}_M = \mathrm{RV}_F$. Flatter and steeper $\mathrm{RV}_M$ are
shown in blue and green, respectively. Note that age 4 is the youngest age at which a female can turn into a male in our OM.

(ref:simpop-plot-fig-caption) Simulated female (left) and male (right) abundance for years 75 to 119 by OM scenario (grey = base, red = skipped spawning, blue = male RC steeper than female, green = male RC flatter than females). Each of the 40 abundance time series per scenario is plotted individually, and the median is presented by bold lines.

(ref:mainckmrpairs-plot-fig-caption) Number of CKMR pairwise relationships for three years of CKMR sampling at $10\sqrt{N}$ by OM scenario. 
$\mathrm{POP}^{\neq}$ = parent-offspring pairs with different mitochondrial DNA, $\mathrm{POP}^=$ = parent-offspring pairs with matching mitochondrial DNA, 
$\mathrm{HSP}^{\neq}$ = half-sibling pairs with different mitochondrial DNA, $\mathrm{HSP}^=$ = half-sibling pairs with matching mitochondrial 
DNA. In the base OM scenario, there is no skipped spawning, male and female reproductive value (RV) are equal. The box represents the interquartile range (IQR), the line is the median, 
the whiskers extend to 1.5 IQR and outliers fall beyond that.

(ref:ckmrsampling-plot-fig-caption) Percent relative error in estimated male and female abundance time series 
summarized across years, by number of CKMR sampling years (red = 3, blue = 10), CKMR sample size, and whether or not sex transition 
($P^{\mathrm{F}\rightarrow \mathrm{M}}$) and male reproductive value 
($\mathrm{RV}_{M}$) functions are were assumed known (fixed). Results are shown for the base OM scenario and without
any model misspecification. Y-axis truncated to improve readability.

(ref:ckmrnll-plot-fig-caption) Percent relative error in estimated male and female abundance time series
summarized across years, by CKMR information availability and whether or not sex transition 
($P^{\mathrm{F}\rightarrow \mathrm{M}}$) and male reproductive value 
($\mathrm{RV}_{M}$) functions are assumed known (fixed). Results are shown for the base CKMR sampling scenario. In the base EM, both half-sibling
pairs and parent-offspring-pairs were used in the likelihood and the assumed mitochondrial haplotype diversity,
$\Hmte$ (set to 1), is equal to the simulated haplotype diversity, $\Hmts$. When $\Hmte \neq \Hmts$, one was set to 1 and the other to 0.9.
Notice the different y-axis scale for panel C. Y-axis truncated to improve readability.

(ref:sexratio-plot-fig-caption) Percent relative error in estimated sex ratio 
(number of estimated males divided by total number of age 3+ year-olds) summarized across years 
under different OM and EM configurations and by whether or not sex transition 
($P^{\mathrm{F}\rightarrow \mathrm{M}}$) and male reproductive value 
($\mathrm{RV}_{M}$) functions are assumed known (fixed). In the Base case there is no skipped spawning, male and female RV are
equal, and $\Hmte = \Hmts = 1$. When $\Hmte \neq \Hmts$, one was set to 1 and the other to 0.9.
Notice the different y-axis scale for panel D. Y-axis truncated to improve readability.

(ref:estfuns-plot-fig-caption) Estimated sex transition and male reproductive value (RV) functions under different OM and EM configurations.
The dashed black lines are the simulated values and solid black lines are the median estimated values. sHmt = simulated mitochondrial haplotype diversity, eHmt = mitochondrial haplotype diversity assumed in the estimation model. In the Base case there is no skipped spawning, male and female RV are equal, and eHmt = sHmt = 1. When eHmt $\neq$ sHmt, one was set to 1 and the other to 0.9.

(ref:malerole-plot-fig-caption) Percent relative error in estimated male and female abundance time series 
summarized across years for the OM scenarios exploring the effect of male reproductive output and by
whether or not sex transition 
($P^{\mathrm{F}\rightarrow \mathrm{M}}$) and male reproductive value 
($\mathrm{RV}_{M}$) functions are assumed known (fixed). When $P^{\mathrm{F}\rightarrow \mathrm{M}}$ was
fixed, it was fixed at the true value. When $\mathrm{RV}_{M}$ was fixed, however, it was fixed at the same
value as that of females. Consequently when $\mathrm{RV}_M$ is fixed (red and blue), it is misspecified in the EM when
$\mathrm{RV}_M^\mathrm{OM}$ differs from $\mathrm{RV}_F^\mathrm{OM}$ (panels C and D). In the base OM, there is no skipped spawning and male
and female RV are equal. Y-axis truncated to improve readability.

\newpage

## Figures

```{r simoverview-plot, echo=FALSE, fig.align='center', fig.height=4, fig.cap = "(ref:simoverview-plot-fig-caption)", message=FALSE, warning=FALSE}
knitr::include_graphics('images/Sim_Overview.pdf')
```

```{r flowchart-plot, echo=FALSE, fig.align='center', fig.height=6.26, fig.cap = "(ref:flowchart-plot-fig-caption)", message=FALSE, warning=FALSE}
knitr::include_graphics('images/ckmr_sims_overview.drawio.pdf')
```

```{r RC-plot, echo=FALSE, fig.align='center', fig.height=3, fig.width=7, fig.cap = "(ref:RC-plot-fig-caption)", message=FALSE, warning=FALSE} 
mRC_plot
```

```{r simpop-plot, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, fig.cap = "(ref:simpop-plot-fig-caption)", message=FALSE, warning=FALSE}
sim_pop_plot
```

```{r mainckmrpairs-plot, echo=FALSE, fig.align='center', fig.height=6, fig.width=8, fig.cap = "(ref:mainckmrpairs-plot-fig-caption)", message=FALSE, warning=FALSE}
main_ckmr_pairs_plot
```

```{r ckmrsampling-plot, echo=FALSE, fig.align='center', fig.height=9, fig.width=9, fig.cap = "(ref:ckmrsampling-plot-fig-caption)", message=FALSE, warning=FALSE}
ckmr_sampling_plot

```

```{r ckmrnll-plot, echo=FALSE, fig.align='center', fig.height=9, fig.width=9, fig.cap = "(ref:ckmrnll-plot-fig-caption)", message=FALSE, warning=FALSE}
ckmr_nll_plot

```

```{r sexratio-plot, echo=FALSE, fig.align='center', fig.height=8, fig.width=8, fig.cap = "(ref:sexratio-plot-fig-caption)", message=FALSE, warning=FALSE}
sr_plot
```

```{r estfuns-plot, echo=FALSE, fig.align='center', fig.height=12, fig.width=11, out.height="500px", fig.cap = "(ref:estfuns-plot-fig-caption)", message=FALSE, warning=FALSE}
est_fun_plot
```

```{r malerole-plot, echo=FALSE, fig.align='center', fig.height=9, fig.width=9, fig.cap = "(ref:malerole-plot-fig-caption)", message=FALSE, warning=FALSE}
male_role_plot

```

\clearpage
\newpage

# Appendix A. Full CKMR Model Specification

\renewcommand{\theequation}{A\arabic{equation}} \setcounter{equation}{0}

To specify the CKMR model, we first derive kinship probabilities representing the 
likelihood that any two individuals in a genetic sample are HSPs or POPs. Then we combine those into
two different CKMR pseudolikelihoods that we tested. See Table \@ref(tab:table1) for a 
summary of notation used in the paper.

### HSP Kinship Probabilities

With protogynous
hermaphrodites, there are three types of HSPs possible, determined by
the sex of the shared parent when each of the half-siblings was born:

- $F\rightarrow F$:  the shared parent was a female ($F$) when producing the first-born and the second-born
members of the HSP.
- $F\rightarrow M$: the shared parent was a female when producing the first-born member and a male ($M$)
when producing the second-born member of the HSP.
- $M\rightarrow M$: The shared parent was a male when producing both the first- and second-born members
of the HSP.

We assume that ages of sampled individuals are known without error, sampling is conducted on dead
individuals, and age-specific survival
is independent of sex. We use an age-structured
model with intervals indexed by year. We furthermore assume the species is monandric (i.e., all males
originate from females) and that sex change, if it occurs, happens at the
beginning of the interval, followed by reproduction, then mortality/fishing/sampling. Thus,
a fish that transitions to male at time $t$ will also be a male when producing offspring
at time $t$ and when sampled in the fishery at time $t$.  The formulas we present here would need
to be adjusted if these events take place in a different order. Let $i$
denote the first-born and $j$ the second-born of a sampled pair, where $c_i \neq c_j$ (cross-cohort comparisons only).  
The half-sibling probabilities for a protogynous
species must include not only the probability of a common parent surviving from the time $i$ was born, $c_i$, to
the time $j$ was born, $c_j$, but also the probability of a female transitioning to male between $c_i$ and $c_j$. 
Define the age difference between $i$ and $j$ as $d=c_{j}-c_{i}$, denote
by $a^*$ the shared
parent’s age at the time $j$ was born, and define $\ERRO_{a^*,t,g}$ as the expected reproductive
output of all the age $a^*$ individuals of sex $g$ ($g \in \{F, M\}$) at time $t$ relative to the population (Table \@ref(tab:table1)):
\begin{equation}
\ERRO_{a^*,t,g} = \frac{N_{a^*,t,g}^*\Mat_{a^*,g}\mathrm{RV}_{a^*,g}}{\sum_{a\in A} N_{a,t,g}^*\Mat_{a,g}\mathrm{RV}_{a,g}},
(\#eq:erro)
\end{equation}
where $\mathrm{Mat}_{a,g}$ and $\mathrm{RV}_{a,g}$ are maturity and reproductive value, respectively, 
of age $a$ and sex $g$ individuals, and $N_{a,t,g}^*$, is the relative abundance of age $a$ and sex $g$ individuals at time $t$.
We define reproductive value at age as the age-specific relative potential to produce offspring. It is typically
associated with mean female fecundity-at-age (more on this below). For males it would correlate with fertilization rates. 
Conditional on $c_i$, $c_j$, and $a^*$, the probability that a randomly sampled pair is an HSP produced by the $F\rightarrow F$ scenario is
\begin{equation}
\begin{aligned}
P(\mathrm{HSP}_{F\rightarrow F}|a^*,c_i, c_j) &= \ERRO_{a^*-d,c_i,F} \\
 &\times \prod_{b=0}^{d-1} S_{a^* - d + b,c_i + b} \\
 &\times \prod_{a = a^*-d+1}^{a^*} (1 - \tprob_{a}) \\
 &\times \frac{1}{N_{a^*,c_j,F}},
\end{aligned}
\end{equation}
where $S_{a,t}$ is the annual age-specific survival probability, $N_{a,t,g}$ is the total abundance of age $a$ and sex $g$ fish at time $t$,
and $\tprob_{a}$ is the probability that an individual that starts age $a$ as a female switches sex to be
an age-$a$ male before reproduction ($b$ is just a summation index).

The above probability arises because, in order for a pair $i$ and $j$ with a mother of age $a^*$ when $j$ was born to form an  $\mathrm{HSP}_{F\rightarrow F}$, the following
four independent events must occur:

1. $i$ must be born to a female of age $a^* - d$. This occurs with probability $\ERRO_{a^*-d,c_i,F}$.
2. The mother of $i$ must survive from year $c_i$ until the time of $j$'s birth, which
happens according to the age-specific survival rates with probability $\prod_{b=0}^{d-1} S_{a^* - d + b,c_i + b}$.
3. The mother of $i$ must remain a female from the time of reproduction at age $a^* -d$ until the time of reproduction at age $a^*$, which
occurs with probability $\prod_{a = a^*-d+1}^{a^*} (1 - \tprob_{a})$.
4. The mother of $j$ must be the same female as the mother of $i$.  Under the
assumption of equality of the expected individual reproductive outputs for fish of the same
sex and age, this probability is simply the reciprocal of the number of
age $a^*$ females at time $c_j$, which can be written as the reciprocal of the
total female abundance at age $a^*$ at time $c_j$: $1/(N_{a^*,c_j,F})$.

It is important to note that $P(\mathrm{HSP}_{F\rightarrow F}|a^*,c_i, c_j)$ is a
probability conditioned on the mother's
age being $a^*$. We obtain the required marginal probability,
$P(\mathrm{HSP}_{F\rightarrow F}|c_i, c_j)$, by summing the conditional
probability over all values of $a^*$, weighted by the probability that $j$'s mother is
of age $a^*$, which is simply $\ERRO_{a^*,c_j,F}$. 
Thus, the marginal probability is:
\begin{equation}
P(\mathrm{HSP}_{F\rightarrow F}|c_i, c_j) = \sum_{d < a^* \leq A} P(\mathrm{HSP}_{F\rightarrow F}|a^*,c_i, c_j)\ERRO_{a^*,c_j,F},
\end{equation}
where $A$ is the maximum age. Note that, in practice, pairs with $c_i=c_j$ are typically excluded, since half siblings from
the same cohort may have correlated juvenile survival which
could lead to downwardly biased estimates of abundance [@waples2022close].

With that first scenario, $F\rightarrow F$, derived, the remaining two follow along similarly,
with differences only in the term describing the probability of changing sex (or not). Thus,
for the scenario $F\rightarrow M$ to occur, the female must transition from being female to
male between times $c_i$ and $c_j$, and the conditional probability is
\begin{equation}
\begin{aligned}
P(\mathrm{HSP}_{F\rightarrow M}|a^*,c_i, c_j) &= \ERRO_{a^*-d,c_i,F} \\
 &\times \prod_{b=0}^{d-1} S_{a^* - d + b,c_i + b} \\
 &\times \biggl[1 - \prod_{a=a^*-d+1}^{a^*} (1 - \tprob_{a})\biggr] \\
 &\times \frac{1}{N_{a^*,c_j,M}}.
\end{aligned}
\end{equation}

The marginal probability of sampling an $\mathrm{HSP}_{F\rightarrow M}$ is then,
\begin{equation}
P(\mathrm{HSP}_{F\rightarrow M}|c_i, c_j) = \sum_{d < a^* \leq A} P(\mathrm{HSP}_{F\rightarrow M}|a^*,c_i, c_j)\ERRO_{a^*,c_j,M}.
\end{equation}

To calculate the probability of encountering an HSP arising from the
$M\rightarrow M$ scenario, there are no sex-transition terms because, in protogynous
hermaphrodites, males remain males until death.  Hence, the conditional probability is
\begin{equation}
\begin{aligned}
P(\mathrm{HSP}_{M\rightarrow M}|a^*,c_i, c_j) &= \ERRO_{a^*-d,c_i,M} \\
 &\times \prod_{b=0}^{d-1} S_{a^* - d + b,c_i + b} \\
 &\times \frac{1}{N_{a^*,c_j,M}},
\end{aligned}
\end{equation}
and the marginal probability is
\begin{equation}
P(\mathrm{HSP}_{M\rightarrow M}|c_i, c_j) = \sum_{a^*\in A} P(\mathrm{HSP}_{M\rightarrow M}|a^*,c_i, c_j)\ERRO_{a^*,c_j,M},
\end{equation}

Because we do not observe the sex of the shared parent at the time of $i$ and $j$'s birth, we
cannot know with certainty whether a half-sibling pair falls into the $F\rightarrow F$,
$F\rightarrow M$, or $M\rightarrow M$, categories; however we can gain partial information
by sequencing the HSP members' mitochondrial DNA (mtDNA), which is inherited from the female parent.
Let $a$ be the haplotype carried by the first-born and $b$ the haplotype carried
by the second born, noting that they may carry the same haplotype ($a=b$) or
different haplotypes ($a\neq b$). We denote the mtDNA haplotype data in the pair $\mathrm{mt}_{a,b}$. Including the mtDNA as observed data,
we then calculate the joint probability
of the kinship/relationship and mtDNA haplotypes of the pair.
In the absence of sequencing errors or mutation during the most recent meioses,
under the $F\rightarrow F$, scenario, both members of
the HSP will have the same mtDNA sequence because the two mtDNAs are recently
identical by descent. By contrast, in the $F\rightarrow M$ and $M\rightarrow M$ scenarios,
the mtDNA of the two HSPs are not identical by descent and can be regarded as two
randomly sampled mtDNA sequences from the population.  Taking $f_a$ and $f_b$ to be the
relative frequency of mtDNA haplotypes carried by the first-born and second-born,
respectively, we have

\begin{equation}
 P(\mathrm{HSP}, \mathrm{mt}_{a,b}|c_i,c_j) = 
  \begin{cases} 
    \begin{aligned} 
      & f_a P(\mathrm{HSP}_{F\rightarrow F}|c_i, c_j) \\ 
      &\quad + f_a^2 \biggl[P(\mathrm{HSP}_{F\rightarrow M}|c_i, c_j)  + P(\mathrm{HSP}_{M\rightarrow M}|c_i, c_j)\biggr],
  \end{aligned} & \mathrm{if}~a=b \\
  \\
  f_a f_b\biggl[P(\mathrm{HSP}_{F\rightarrow M}|c_i, c_j) + P(\mathrm{HSP}_{M\rightarrow M}|c_i, c_j)\biggr], & \mathrm{if}~a \neq b.
  \end{cases}
  \label{eq:hspmtf}
\end{equation}

### POP Kinship Probabilities

For parent-offspring pair (POP) probabilities, we assume that the sex of sampled individuals
can be determined without error and that samples come from the fishery and are
caught after reproduction in any give year. (Note that knowledge about sampled individuals'
sex is not required if only HSPs are used in the likelihood.)
Since we assume sampling is lethal, the sampling year of the first-born individual in a pair,
$i$, (i.e., the potential parent) must be the same or later than the birth year
of the second-born individual, $j$ (the potential offspring) in order for a pair to be a POP.
The sex of the candidate parent must
be accounted for in the kinship probability, and it is also necessary to
account for the possibility that a female parent changes sex to become
a male before being sampled.

We let $s_i$ denote the sampling year of $i$.  If $i$ is sampled in the same
year as $j$'s birth ($s_i = c_j$),
then $i$ could not have changed sex after producing $j$ (recall that we assume
sex-change occurs each year before reproduction), and the probability of a POP
can be calculated by the standard approach. Taking first the case when
$i$ is a female, a pair
$(i,j)$ with $c_i < c_j$ and $s_i = c_j$ will be a
$\mathrm{POP}$ if the following two events occur:

1. $j$ must be born to a female
that is the same age as $i$ at time $c_j$. At time $c_j$,
$i$ will be of age $d = c_j - c_i$, and the probability of being produced by an
age-$d$ female in year $c_j$ is the expected relative reproductive output,
$\ERRO_{d,c_j,F}$.
2. From amongst the $N_{d,c_j,F}$ age-$d$ females producing offspring
at time $c_j$, the mother of $j$ must be just one of them, $i$, which occurs with
probability equal to $1/(N_{d,c_j,F})$.

Thus, we have, if $i$ is a female when sampled at time $c_j$,
\begin{equation}
P(\mathrm{POP}|c_i, c_j, g_i=F,s_i=c_j) = \frac{\ERRO_{d,c_j,F}}{N_{d,c_j,F}},
\end{equation}
where $g_i=F$ indicates that $i$ was of sex $F$ at the time of sampling.  The above
translates easily to the situation when $i$ is a male:
\begin{equation}
P(\mathrm{POP}|c_i, c_j, g_i=M,s_i=c_j) = \frac{\ERRO_{d,c_j,M}}{N_{d,c_j,M}}.
\end{equation}
It is worth noting that, if we write the expected total reproductive output of sex-$g$ individuals
at time $t$ as $\ETRO_{t,g} = \sum_{a\in A} N_{a,t,g}\Mat_{a,g}\mathrm{RV}_{a,g}$ (this is
simply the product of $N_{t,g}$ and the denominator in \@ref(eq:erro)), we can rewrite these probabilities as:
\begin{equation}
P(\mathrm{POP}|c_i, c_j, g_i,s_i=c_j) =  \frac{\mathrm{RV}_{d,g_i} \mathrm{Mat}_{d,g_i}}{\ETRO_{c_j,g_i} }
(\#eq:popsimple)
\end{equation}
for individual $i$ of sex $g_i$, with  $\mathrm{Mat}_{d,M} \equiv 1$ because in protogynous hermaphrodites,
all males are mature.

When sampling of $i$ occurs a year or more after the birth of $j$, it is necessary
to account for the possibility that the parent changed sex after giving birth to
$j$ but before being sampled. If $i$ is a male at time $c_j$, then it will also
be male at all time points in the future.  This POP scenario is denoted by $\mathrm{POP}_{M\rightarrow M}$
and the probability of its occurrence requires no correction due to changing sex, and
thus we can combine this with the result in \@ref(eq:popsimple) for males, giving:
\begin{equation}
P(\mathrm{POP}_{M\rightarrow M}|c_i, c_j, s_i, g_{i,s_i} = M) = \mathbbold{1}\{s_i \geq c_j\} \frac{\mathrm{RV}_{d,c_j,M} \mathrm{Mat}_{d,c_j,M}}{\ETRO_{c_j,M} },
\end{equation}
where $\mathbbold{1}\{s_i \geq c_j\}$ is an indicator function that evaluates to 1
if $i$ was sampled after $j$'s birth, and 0 otherwise, and where we added the $s_i$ subscript to $g_i$ to emphasize that $g_{i, s_i}$ is referring to the sex of the
candidate parent at the time of its sampling at $s_i$.

If $i$ was female at time $c_j$, but sampled at time $s_i > c_j$, then there is a
chance that $i$ will have changed sex.  We denote the two possibilities as
$\mathrm{POP}_{F\rightarrow F}$ for the case with no sex change, and $\mathrm{POP}_{F\rightarrow M}$
for the case when sex change occurs. The probabilities for these cases with $s_i > c_j$ are
\begin{equation}
\begin{aligned}
P(\mathrm{POP}_{F\rightarrow F}|c_i, c_j, s_i, g_{i,s_i} = F) &= \mathbbold{1}\{s_i \geq c_j\} \frac{\mathrm{RV}_{d,c_j,F} \mathrm{Mat}_{d,c_j,F}}{\ETRO_{c_j,F}} 
\prod_{d< a \leq s_i - c_i} (1 - \tprob_{a})
\\
P(\mathrm{POP}_{F\rightarrow M}|c_i, c_j, s_i, g_{i,s_i} = M) &= \mathbbold{1}\{s_i \geq c_j\} \frac{\mathrm{RV}_{d,c_j,F} \mathrm{Mat}_{d,c_j,F}}{\ETRO_{c_j,F} } 
\biggl[ 1 - \prod_{d< a \leq s_i - c_i} (1 - \tprob_{a})\biggr].
\end{aligned}
\end{equation}

Of the three above probabilities, only $P(\mathrm{POP}_{F\rightarrow F}|c_i, c_j, s_i)$ can be
used directly when $s_i > c_j$, because we do not observe the sex of $i$ at time $c_j$, and, when
$i$ is a male at time $s_i > c_j$, $i$ could have been either male or female at time $c_j$; however,
as before, we can garner some information from mtDNA.  Thus, similar to the case for HSPs,
conditional on $g_i$,
we calculate the joint probability that the pair
is a POP and that the parent and the offspring carry mtDNA sequence $a$ 
and $b$, respectively.

\begin{equation}
P(\mathrm{POP}, \mathrm{mt}_{a,b} | c_i, c_j, g_i, s_i)=
\begin{cases}
f_a P(\mathrm{POP}_{F\rightarrow F}|c_i, c_j, s_i), & \mathrm{if}~a=b,~g_{i,s_i}=F \\
0,                                                  & \mathrm{if}~a\neq b,~g_{i,s_i}=F \\
\begin{aligned} &f_a P(\mathrm{POP}_{F\rightarrow M}|c_i, c_j, s_i)  \\
&   \quad  + f_a^2 P(\mathrm{POP}_{M\rightarrow M}|c_i, c_j, s_i),
\end{aligned} & \mathrm{if}~a=b,~g_{i,s_i}=M \\
 f_a f_b P(\mathrm{POP}_{M\rightarrow M}|c_i, c_j, s_i), & \mathrm{if}~a\neq b,~g_{i,s_i}=M
\end{cases}
(\#eq:popcases)
\end{equation}

The probabilities for observing
POPs \@ref(eq:popcases),  or HSPs \@ref(eq:hspmtf),
along with their mtDNA haplotypes 
form the terms in the CKMR pseudolikelihood.

### The CKMR Pseudolikelihood

The pairs of CKMR samples in the data can be summarized into the counts of sample pairs
in different groups according to their covariates: those
born in different years, $c_i < c_j$, with the first-born having sex $g_i \in \{F,M\}$ when
lethally sampled at time $s_i$.   These pairs summarized by $c_i, c_j, g_i, s_i$ may be
further distributed into categories according to their
kin relationship and mtDNA haplotypes. We let $\K{HSP}{c_i,c_j}{g_i,s_i}{a,b}$
$\K{POP}{c_i,c_j}{g_i,s_i}{a,b}$, and $\K{Not}{c_i,c_j}{g_i,s_i}{a,b}$
denote the number of HSPs or POPs or pairs that are neither HSPs or POPs, respectively,
observed among the sample pairs with
a first-born from year $c_i$ sampled as sex $g_i$ in year $s_i$ carrying a mtDNA haplotype
denoted by $a$, and a second-born from $c_j$ carring a mtDNA haplotype denoted $b$.

Following the standard pseudolikelihood approach for
CKMR (Bravington et al. 2016), we assume that all sample pairs can be treated independently and thus the
joint probability of all the sample counts
for any birth years $c_i < c_j$, sampling years $s_i$, first-born sex at sampling $g_i$, and mtDNA
haplotypes $a$ and $b$ is
approximated by a product of multinomial probabilities. This provides the CKMR likelihood
including both HSPs and POPs:
\begin{equation}
\begin{aligned}
L_\mathrm{HP} \propto \prod_{ {c_i < c_j \atop g_i \in \{F, M\}} \atop Y_\mathrm{lo} \leq s_i \leq Y_\mathrm{lo}}
\prod_{ {a \in \sH \atop b \in \sH}}
& P(\mathrm{HSP}, \mathrm{mt}_{a,b}|c_i,c_j) ^ {\K{HSP}{c_i,c_j}{g_i,s_i}{a,b}} \\
& \times P(\mathrm{POP}, \mathrm{mt}_{a,b} | c_i, c_j, g_i, s_i)^{\K{POP}{c_i,c_j}{g_i,s_i}{a,b}} \\
& \times P(\mathrm{Not}, \mathrm{mt}_{a,b} | c_i, c_j, g_i, s_i)^{\K{Not}{c_i,c_j}{g_i,s_i}{a,b}}
\end{aligned}
\end{equation}
where $Y_\mathrm{lo}$ and $Y_\mathrm{hi}$ are the first and last years, respectively,
when CKMR samples are taken from the fishery, $\sH$ is the set of all known mtDNA haplotypes, and
\begin{equation}
\begin{aligned}
P(\mathrm{Not}, \mathrm{mt}_{a,b} | c_i, c_j, g_i, s_i) = 1 - P(\mathrm{HSP}, \mathrm{mt}_{a,b}|c_i,c_j) - P(\mathrm{POP}, \mathrm{mt}_{a,b} | c_i, c_j, g_i, s_i).
\end{aligned}
\end{equation}
It is worth emphasizing that $\K{POP}{c_i,c_j}{g_i,s_i}{a,b}$ must be
zero for all values of $s_i < c_j$.
For comparison purposes, one may also formulate the likelihood including only HSPs (and ignoring POPs), as follows:
\begin{equation}
\begin{aligned}
L_\mathrm{H} \propto \prod_{ {c_i < c_j \atop \mbox{}}}
\prod_{ {a \in \sH \atop b \in \sH}}
& P(\mathrm{HSP}, \mathrm{mt}_{a,b}|c_i,c_j) ^ {\K{HSP}{c_i,c_j}{}{a,b}} \\
& \times [1 - P(\mathrm{HSP}, \mathrm{mt}_{a,b}|c_i,c_j)]^{\mathring{K}^{\mathrm{Not},a,b}_{c_i,c_j}}
\end{aligned}
\end{equation}
where $\KHSs{c_i,c_j}{a,b} = \sum_{g_i,s_i} \K{HSP}{c_i,c_j}{g_i,s_i}{a,b}$ and $\mathring{K}^\mathrm{Not,a,b}_{c_i,c_j}$ denotes all pairs with birth years $c_i$ and $c_j$ and
mtDNA haplotypes $a$ and $b$ that are not HSPs.

# Appendix B. Estimation Model Objective Function

\renewcommand{\theequation}{B\arabic{equation}} \setcounter{equation}{0}

The joint total objective function, $L_T$, for the observed CKMR and fishery data was the
sum of the weighted CKMR, recruitment index, and age composition components and a 
penalty vector, $L_P$:

\begin{equation}
L_T={\omega_I}L_I + {\omega_C}L_C + {\omega_{HP}}L_{HP} + L_P,
\end{equation}

The likelihood for the index data, $L_I$, was
calculated following the approach described by Walters and Ludwig (1994), by assuming that
the differences between the observed and predicted abundance index
were normally distributed and using the maximum likelihood estimate for survey
catchability $q$. The index log-likelihood (without the constant terms) was thus:

\begin{equation}
L_I = \sum \log(\sigma_I) + (Y_t-\frac{1}{n} \sum Y_t)^2 / 2 \sigma_I^2,
\end{equation}

where $Y_t$ was calculated as:

\begin{equation}
Y_t = log(I_t) - log(\hat{I_t}).
\end{equation}

The composition data likelihood, $L_C$, was specified using a multivariate logistic 
likelihood (Richards et al. 1997). This likelihood, implemented via ADMB's *dmvlogistic* 
function, accounts for differences between observed and expected proportions 
while incorporating overdispersion effects. See main text for the CKMR likelihood, $L_{HP}$.

The penalty vector, $L_P$, consisted of lognormal penalties for the mean fishing mortality
parameters, $\bar{F}_r$ and $\bar{F}_c$, (with $\mu_F$ = -1.2 $\sigma_F$ = 0.8) and the sex
transition function $SD$ parameter, $\mathrm{STF}_{SD}$ 
(with $\mu_{STF}$ = 1.5 $\sigma_{STF}$ = 0.5) 
as well as penalties for recruitment ($P_{\mathrm{Rdev}}$) and fishing mortality ($P_{\mathrm{Fdev}}$) deviations:

\begin{equation}
L_P = P_{\bar{F}_r} + P_{\bar{F}_c} + P_{STF_{SD}} + P_{\mathrm{Rdev}} + P_{\mathrm{Fdev_r}} + P_{\mathrm{Fdev_c}}.
\end{equation}

The penalties on deviations were set to be weak normal ($\sigma_{dev}$ = 2) until the last 
phase of estimation when they were set as 100 times the sum of the squared elements.

The initial CKMR and index data weights, $\omega$, were set to one, while the initial composition 
data weight was specified as the number of abundance
index data points (equal to the number of assessment years with data points) 
divided by the age composition data points (equal to number of assessment years times number
of ages modeled, if no missing data). After the initial run, fit statistics were
calculated for each data source and weights were adjusted following the recommendations
of Francis (2011) and Francis (2017). Abundance index weights were
only allowed to increase and CKMR data weights were only allowed to decrease,
while composition data weights could increase or decrease. The model was
refitted up to two times for each iteration. If the model was run more than
once, the run with the highest likelihood was retained for subsequent analysis.

## References

\begin{hangparas}{.25in}{1}

Bravington, M.V., Skaug, H.J., and Anderson, E.C. 2016b. Close-kin mark-recapture.
Statistical Science 31(2): 259–274.

Francis, R.C. 2011. Data weighting in statistical fisheries stock assessment models.
Canadian Journal of Fisheries and Aquatic Sciences $\textbf{68}$(6): 1124–1138.

Francis, R.C. 2017. Revisiting data weighting in fisheries stock assessment models.
Fisheries Research $\textbf{192}$: 5–15.

Richards, L.J., Schnute, J.T., and Olsen, N. 1997. Visualizing catch–age analysis: A case
study. Canadian Journal of Fisheries and Aquatic Sciences $\textbf{54}$(7): 1646–1658. NRC
Research Press Ottawa, Canada.

Walters, C., and Ludwig, D. 1994. Calculation of Bayes posterior probability distributions
for key population parameters. Canadian Journal of Fisheries and Aquatic Sciences $\textbf{51}$(3): 713–722.

Waples, R.S., and Feutry, P. 2022. Close-kin methods to estimate census size and effective
population size. Fish and Fisheries 23(2): 273–293.
\end{hangparas}