### 2025-05-14 manuscript/fig_captions.Rmd
change manuscript format to Fisheries Research

---
title             : "Application of Close-kin mark–recapture (CKMR) to Protogynous Fishes"
subtitle          : "Figure Captions"
author            : 
  - Claudia Friess, Eric C. Anderson, Lisa E. Ailloud, David S. Portnoy, and
  - Susan K. Lowerre-Barbieri

bibliography      : ["citations_abbrev.bib", "r-references.bib"]
csl               : "elsevier-harvard.csl"

floatsintext      : yes
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no
always_allow_html : true

documentclass     : article
output            : pdf_document
geometry          : margin=1in
linestretch       : 1.5
fontsize          : 12pt
header-includes   : 
  - \raggedright 
  - \usepackage{color}
  - \usepackage{tipa}
  - \usepackage{wasysym}
  - \usepackage{fancyhdr}
  - \usepackage{mathrsfs}
  - \thispagestyle{empty}
  - \usepackage{hanging}
  - \usepackage{setspace}
  - \usepackage{caption} 
  - \setlength{\baselineskip}{1.5\baselineskip}
  - \newcommand{\tprob}{P^{F\rightarrow M}}
  - \newcommand{\wt}[1]{\omega_#1}  % recruitment deviations
  - \newcommand{\rbar}{\bar{R}}  % mean recruitment
  - \newcommand{\fbarr}{{\bar{F}_\mathrm{rec}}}  % mean fishing mortality from rec fishery
  - \newcommand{\frecdev}[1]{fdev^\mathrm{rec}_#1}  % annual deviations from mean mortality from rec fishery
  - \newcommand{\fbarc}{{\bar{F}_\mathrm{com}}}  % mean fishing mortality from commercial fishery
  - \newcommand{\fcomdev}[1]{fdev^\mathrm{com}_#1}  % annual deviations from mean mortality from rec fishery
  - \newcommand{\Mage}[1]{\dot{M}_#1}  % age-specific instantaneous natural mortality
  - \newcommand{\vulr}[1]{\dot{v}^\mathrm{rec}_#1}  % age-specific vulnerability to recreational fishery
  - \newcommand{\vulc}[1]{\dot{v}^\mathrm{com}_#1}  % age-specific vulnerability to commercial fishery
  - \newcommand{\dFec}[1]{\dot{Fec}_#1}    % age-specific relative fecundity
  - \newcommand{\Af}[1]{\dot{Mat}_#1}    % age-specific fraction of females mature
  - \newcommand{\Pfm}[1]{\dot{P}^{\mathrm{F}\rightarrow\mathrm{M}}_#1}    % age-specific prob of transitioning from female to male
  - \newcommand{\Imp}[1]{\dot{P}^\mathrm{M}_#1}    % equilibrium proportion of males amongst age a fish
  - \newcommand{\It}[1]{\mathring{I}_{#1}}   % recruitment index in year t
  - \newcommand{\pacom}[2]{\mathring{p}^\mathrm{com}_{#1,#2}}   % relative abundance of age a females in year t
  - \newcommand{\pacr}[3]{\mathring{p}^\mathrm{ckmr}_{#1,#2,#3}}   % relative abundance of age a males in year t
  - \newcommand{\Z}[2]{Z_{#1,#2}}   % total instantaneous mortality at time t for age a fish 
  - \newcommand{\AS}[2]{S_{#1,#2}}   % total annual survival
  - \newcommand{\lxf}[1]{s_{#1}}     % survivorship
  - \newcommand{\Nf}[2]{N^\mathrm{fem}_{{#1}, {#2}}}  % number of females of age a at time t
  - \newcommand{\Nm}[2]{N^\mathrm{mal}_{{#1}, {#2}}}  % number of males of age a at time t
  - \newcommand{\K}[4]{\mathring{K}^{\mathrm{#1},{{#4}}}_{{#2},{#3}}}
  - \newcommand{\KHS}[3]{\mathring{K}^{\mathrm{HSP},{{#3}}}_{{#1},{#2}}}
  - \newcommand{\KHSs}[2]{\mathring{K}^{\mathrm{HSP},{{#2}}}_{{#1}}}
  - \newcommand{\KPOPT}[3]{\mathring{K}^{\mathrm{POP},{{#3}}}_{{#1},{#2}}}
  - \newcommand{\KHSD}[2]{\mathring{K}^{\mathrm{HSP}^{\neq _{a,b}}}_{{#1},{#2}}}   % number of half-sibling pairs (born c1 and c2) with different mtDNA 
  - \newcommand{\KHSS}[2]{\mathring{K}^{\mathrm{HSP}^{=_a}}_{{#1},{#2}}}   % number of half-sibling pairs (born c1 and c2) with same mtDNA
  - \newcommand{\KPOP}[3]{\mathring{K}^{\mathrm{POP}^{#3}}_{{#1},{#2}}}  
  - \newcommand{\KPOPF}[2]{\mathring{K}^\mathrm{POP_F}_{{#1},{#2}}}   % number of half-sibling pairs (born c1 and c2) with same mtDNA
  - \newcommand{\KPOPM}[2]{\mathring{K}^\mathrm{POP_M}_{{#1},{#2}}}   % number of half-sibling pairs (born c1 and c2) with same mtDNA 
  - \newcommand{\HSPD}{\mathrm{HSP}^{\neq}}
  - \newcommand{\HSPS}{\mathrm{HSP}^=}
  - \newcommand{\POPD}{\mathrm{POP}^{\neq}}
  - \newcommand{\POPS}{\mathrm{POP}^=}
  - \newcommand{\KNot}[2]{\mathring{K}^\mathrm{Not}_{{#1},{#2}}}     % number of pairs that are neither HSD or HSS
  - \newcommand{\KNote}[2]{\mathring{K}^{\mathrm{Not}=_a}_{{#1},{#2}}}
  - \newcommand{\KNotu}[2]{\mathring{K}^{\mathrm{Not}{\neq}_{a,b}}_{{#1},{#2}}}   
  - \newcommand{\Hmt}{H_\mathrm{mt}}  % mtDNA haplotype diversity
  - \newcommand{\Hmts}{H_\mathrm{mt}^{\mathrm{OM}}}  % simulated mtDNA haplotype diversity
  - \newcommand{\Hmte}{H_\mathrm{mt}^{\mathrm{EM}}}  % mtDNA haplotype diversity assumed in the EM
  - \newcommand{\Mat}{\mathrm{Mat}}
  - \newcommand{\Fec}{\mathrm{Fec}}
  - \DeclareMathAlphabet{\mathbbold}{U}{bbold}{m}{n}
  - \newcommand{\comment}[1]{\textcolor{red}{#1}}
  - \newcommand{\ERRO}{\mathrm{ERRO}}
  - \newcommand{\ETRO}{\mathrm{ETRO}}
  - \newcommand{\sH}{\mathscr{H}}
---

**Figure 1.** Schematic overview of the simulation routine. The
operating model (OM) concerns the simulation of individuals in the population, with four
different scenarios listed. Observed kin pair data are obtained by sampling (sampling model) from the simulated population with different scenarios varying by the number of CKMR sampling years and 
number of samples collected. Finally, several estimation model (EM) configurations were specified. Estimated abundance and sex ratio 
were compared to simulated values by calculating percent relative error. The scenarios in each submodel highlighted by bold font are the base case.
RV = reproductive value, P(\male $\rightarrow$ \female) = sex transition probability. $\Hmts$ = simulated mitochondrial 
haplotype diversity, $\Hmte$ = mitochondrial haplotype diversity assumed in the estimation model.

\vspace{1em} 
**Figure 2.** Overview of annual steps simulated in the individual-based operating model.

\vspace{1em} 
**Figure 3.** Simulated A) reproductive value (RC) at age and B) probability of skipping spawning as a function of sex ratio. In A), the dashed line corresponds to relative female fecundity at age. The solid lines are male RC, where red is the OM scenario where $\mathrm{RV}_M = \mathrm{RV}_F$. Flatter and steeper $\mathrm{RV}_M$ are
shown in blue and green, respectively. Note that age 4 is the youngest age at which a female can turn into a male in our OM.

\vspace{1em} 
**Figure 4.** Simulated female (left) and male (right) abundance for years 75 to 119 by OM scenario (grey = base, red = skipped spawning, blue = male RC steeper than female, green = male RC flatter than females). Each of the 40 abundance time series per scenario is plotted individually, and the median is presented by bold lines.

\vspace{1em} 
**Figure 5.** Number of CKMR pairwise relationships for three years of CKMR sampling at $10\sqrt{N}$ by OM scenario. 
$\mathrm{POP}^{\neq}$ = parent-offspring pairs with different mitochondrial DNA, $\mathrm{POP}^=$ = parent-offspring pairs with matching mitochondrial DNA, 
$\mathrm{HSP}^{\neq}$ = half-sibling pairs with different mitochondrial DNA, $\mathrm{HSP}^=$ = half-sibling pairs with matching mitochondrial 
DNA. In the base OM scenario, there is no skipped spawning, male and female reproductive value (RV) are equal. The box represents the interquartile range (IQR), the line is the median, 
the whiskers extend to 1.5 IQR and outliers fall beyond that.

\vspace{1em} 
**Figure 6.** Percent relative error in estimated male and female abundance time series 
summarized across years, by number of CKMR sampling years (red = 3, blue = 10), CKMR sample size, and whether or not sex transition 
($P^{\mathrm{F}\rightarrow \mathrm{M}}$) and male reproductive value 
($\mathrm{RV}_{M}$) functions are were assumed known (fixed). Results are shown for the base OM scenario and without
any model misspecification. Y-axis truncated to improve readability.

\vspace{1em} 
**Figure 7.** Percent relative error in estimated male and female abundance time series
summarized across years, by CKMR information availability and whether or not sex transition 
($P^{\mathrm{F}\rightarrow \mathrm{M}}$) and male reproductive value 
($\mathrm{RV}_{M}$) functions are assumed known (fixed). Results are shown for the base CKMR sampling scenario. In the base EM, both half-sibling
pairs and parent-offspring-pairs were used in the likelihood and the assumed mitochondrial haplotype diversity,
$\Hmte$ (set to 1), is equal to the simulated haplotype diversity, $\Hmts$. When $\Hmte \neq \Hmts$, one was set to 1 and the other to 0.9.
Notice the different y-axis scale for panel C. Y-axis truncated to improve readability.

\vspace{1em} 
**Figure 8.** Percent relative error in estimated sex ratio 
(number of estimated males divided by total number of age 3+ year-olds) summarized across years 
under different OM and EM configurations and by whether or not sex transition 
($P^{\mathrm{F}\rightarrow \mathrm{M}}$) and male reproductive value 
($\mathrm{RV}_{M}$) functions are assumed known (fixed). In the Base case there is no skipped spawning, male and female RV are
equal, and $\Hmte = \Hmts = 1$. When $\Hmte \neq \Hmts$, one was set to 1 and the other to 0.9.
Notice the different y-axis scale for panel D. Y-axis truncated to improve readability.

\vspace{1em} 
**Figure 9.** Estimated sex transition and male reproductive value (RV) functions under different OM and EM configurations.
The dashed black lines are the simulated values and solid black lines are the median estimated values. sHmt = simulated mitochondrial haplotype diversity, eHmt = mitochondrial haplotype diversity assumed in the estimation model. In the Base case there is no skipped spawning, male and female RV are equal, and eHmt = sHmt = 1. When eHmt $\neq$ sHmt, one was set to 1 and the other to 0.9.

\vspace{1em} 
**Figure 10.** Percent relative error in estimated male and female abundance time series 
summarized across years for the OM scenarios exploring the effect of male reproductive output and by
whether or not sex transition 
($P^{\mathrm{F}\rightarrow \mathrm{M}}$) and male reproductive value 
($\mathrm{RV}_{M}$) functions are assumed known (fixed). When $P^{\mathrm{F}\rightarrow \mathrm{M}}$ was
fixed, it was fixed at the true value. When $\mathrm{RV}_{M}$ was fixed, however, it was fixed at the same
value as that of females. Consequently when $\mathrm{RV}_M$ is fixed (red and blue), it is misspecified in the EM when
$\mathrm{RV}_M^\mathrm{OM}$ differs from $\mathrm{RV}_F^\mathrm{OM}$ (panels C and D). In the base OM, there is no skipped spawning and male
and female RV are equal. Y-axis truncated to improve readability.